### 2016-02-04 misc/labmtg.rmd
lab mtg stuff

---
title: HIV virulence evolution in structured epidemic models
author: Ben Bolker and Daniel Park
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
bibliography: "../manuscript1/virulence.bib"
csl: apa.csl
output:
  ioslides_presentation
---
<!-- 
apa.csl is a slightly hacked version of APA 
  (modified for "et al" after 2 authors in text)
-->
<!-- .refs is style for reference page (small text) -->
<style>
.refs {
   font-size: 12px;
}
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
.title-slide {
   background-color: #55bbff;
}
div#before-column p.forceBreak {
	break-before: column;
}

div#after-column p.forceBreak {
	break-after: column;
}
</style>
<!--    content: url(https://i.creativecommons.org/l/by-sa/4.0/88x31.png)
>
<!-- Limit image width and height -->
<style type="text/css">
img {     
  max-height: 560px;     
  max-width: 800px; 
}
</style>

```{r setup,echo=FALSE,message=FALSE}
library("ggplot2"); theme_set(theme_classic())
library("MASS")
scale_colour_discrete <- function(...,palette="Set1") scale_colour_brewer(...,palette=palette,guide=guide_legend(reverse=TRUE))
scale_fill_discrete <- function(...,palette="Set1") scale_fill_brewer(...,palette=palette,guide=guide_legend(reverse=TRUE))
library("gridExtra")
library("plyr")  ## for ldply()
library("dplyr") ## for full_join()
library("reshape2")  ## for melt()
library("GGally")
## devtools::install_github("lionel-/ggstance")
library("ggstance")
library("knitr")
opts_chunk$set(echo=FALSE,fig.width=4,fig.height=4) ## ,out.width=400)
```

# Intro to virulence evolution

## Tradeoff theory

- assume that virulence evolution is mediated by parasite-level tradeoff between transmission and "virulence" (clearance/mortality rate)
- need a *saturating* relationship: $\frac{d^2 \, \textrm{transmission}}{d\,\textrm{clearance}^2}<0$
- mediated through pathogen replication rate (?)
- simple form assumes no within-host competition (single clone per host)
- alternatives: short-sighted evolution, within-host competition, effects of genetic drift ... [@LevinBull1994,@EbertBull2003]

## Eco-evolutionary dynamics

- feedback between ecological and evolutionary processes
- typical emphasis on *short-term* dynamics (cf. adaptive dynamics)
- eco-evo-epi: *transient virulence* [@bolker_transient_2010]
- balance between $r$-selection (epidemic phase) and ${\cal R}_0$-selection (endemic phase)
- transient peak observed in phage systems [@berngruber_evolution_2013]

## Virulence evolution in HIV

- set-point viral load (SPVL)
- @Fraser+2007 : plausible tradeoff observed from Rakai data
    - heritability of SPVL
	- correlated with rate of progression to AIDS
	- correlated with probability (rate) of within-couple transmission
- @shirreff_transmission_2011 : transient peak after $\approx 1$ in a *reasonably* realistic model

# Modeling virulence evolution in HIV

## @shirreff_transmission_2011 : evolutionary/within-host

- generally based on Rakai estimates
- Normally distributed mutation
- Normally distributed phenotypic variation
- different infectivity for early, mid, late HIV stages
- Hill functions ($x^d/(a+bx^d)$) for infectivity, progression through mid stage
- Weibull time in mid stage

## @shirreff_transmission_2011 : epidemiological

- **very simple**
- implicit partnership-formation model
$$
\textrm{foi} = \frac{\beta c}{\beta + c + d}
$$

- derived from ${\cal R}_0$ for an instantaneous-partnership-formation model, e.g. @hollingsworth_hiv1_2008, @diekmann_mathematical_2000
- (@herbeck_hiv_2014 is worse [I think])

## Epidemiological models for HIV [@champredon_hiv_2013]

<!-- convert journal.pone.0082906.g001.png -crop 100%x50% output.png-->

![](champ1.png)

## Model modifications

- @champredon_hiv_2013 track $\{X,Y,N,D,P\}$ ($\{S,I,SS,SI,II\}$) $\times$ infected strain
- we need $\{S,SS,I(\alpha),SI(\alpha),II(\alpha,\alpha')\}$
- simplifications:
    - single-stage, exponential infectious period  
($\beta =$ time-weighted average of $\beta_1$, $\beta_2(\alpha)$, $\beta_3$)
	- no phenotypic variation

## Model variations

- 4 models:
    - extra-pair contact (yes/no) 
    - instantaneous partnership formation (yes/no)
- implicit (Shirreff/Hollingsworth) model

## Latin hypercube sampling {.columns-2}

- sample evenly, randomly across (potentially many) uncertain parameters
- following @champredon_hiv_2013, earlier @blower_drugs_1991

<p class="forceBreak"></p>

![](LHScrop.png)
[User:Saittam, Wikipedia](https://commons.wikimedia.org/wiki/File:LHSsampling.png#/media/File:LHSsampling.png)

## Details

- $r$ varies widely across parameters/model sets
     - *scale* parameters to $r=0.04$ (partnership formation, dissolution, *and* transmission probability)
     - scale both uncoupled and extra-coupled rates relative to within-coupled rate
- *response variables*
     - equilibrium virulence
	 - peak virulence time (years)
	 - peak virulence
	 - relative peak virulence (peak/eq)

# Results

```{r load,echo=FALSE,cache=TRUE,message=FALSE}
## get all the batch output and combine it into a convenient (?) format
source("../hivFuns.R")  ## must be loaded before LHSparms3
source("../LHSparms3.R")
ltab <- data.frame(run=seq(nrow(ltab)),ltab)
##  combine them/rename appropriately
run_nums <- c("pairform+epc"=5,"pairform"=6,
               "instswitch"=7,"instswitch+epc"=8,"implicit"=9)
## each entry has
##    I_mat*:
##    vir_mat*:
##    eq_vec*: equilibrium value
##    peak_mat*: peak value (time, height)
##    val_vec*: ???
for (i in run_nums)
    L <- load(paste0("ev_LHS_res",i,".rda"))
## combine
component_names <- c("I_mat","vir_mat","eq_vec","peak_mat","val_vec")
## make one big nested list of components
comp_list <- lapply(component_names,
                    function(c) {
    res <- lapply(paste0(c,run_nums),get)
    setNames(res,names(run_nums))
})
names(comp_list) <- component_names
mat_names <- c("I_mat","vir_mat")
## reduce matrices to melted data frame of mean & quantiles
##  (we don't really need to make giant pictures with all of the
##   individual lines - ribbons with mean and 95% quantiles should
##   be sufficient)
sum_mat_fun <- function(x) {
    data.frame(
        tvec=1:nrow(x),  ## ?? better time index?
        mean=rowMeans(x),
        lwr=apply(x,1,quantile,0.025),
        upr=apply(x,1,quantile,0.975))
}
sum_list <- list()
for (m in mat_names) {
    sum_list[[m]] <- ldply(comp_list[[m]],
                           sum_mat_fun, .id="model")
}
s1 <- ldply(comp_list[["eq_vec"]],
                function(x) data.frame(run=1:length(x),eq_vir=x),
                .id="model")
s2 <- ldply(comp_list[["peak_mat"]],
                function(x) {
                   x <- setNames(data.frame(x),c("peak_time","peak_vir"))
                   data.frame(run=1:nrow(x),x)
                 },
                .id="model")
sum_list[["sum_mat"]] <- full_join(s1,s2)
## unlike merge(), full_join() preserves run order
```

```{r fix_order}
mods <- c("pairform+epc","instswitch+epc","instswitch","pairform","implicit")
fixfac <- function(x) {
    mutate(x,model=factor(model,levels=mods))
}
sum_list[["vir_mat"]] <- fixfac(sum_list[["vir_mat"]])
sum_list[["I_mat"]] <-   fixfac(sum_list[["I_mat"]])
```

## Epidemic trajectories

```{r epi_traj,fig.width=10}
gg_Itraj <- ggplot(sum_list[["I_mat"]],
       aes(tvec,mean,ymin=lwr,ymax=upr))+
    geom_line(aes(colour=model))+
    geom_ribbon(aes(fill=model),alpha=0.4)+
    labs(y="mean prevalence")
grid.arrange(gg_Itraj+ggtitle("linear scale"),
             gg_Itraj+scale_y_log10()+ggtitle("log scale"),nrow=1)
```

## Virulence trajectories {.centered}

```{r vir_traj,fig.width=6}
(gg_virtraj <- ggplot(sum_list[["vir_mat"]],
       aes(tvec,mean,ymin=lwr,ymax=upr))+
    geom_line(aes(colour=model))+
    geom_ribbon(aes(fill=model),alpha=0.4))
```

## Univariate distributions

```{r unidist,fig.width=7}
sL <- fixfac(transform(sum_list[["sum_mat"]],rel_peak=peak_vir/eq_vir))
mL <- fixfac(melt(sL,id.vars=c("model","run")))
ggplot(mL,aes(value,model,fill=model))+geom_violinh()+
    facet_wrap(~variable,scale="free")+
    labs(y="",x="")
```

## Pairs plots {.centered}

```{r ggpairs_ex,fig.width=7,fig.height=6.5,dev="png"}
ggp1 <- ggpairs(sL,
        mapping = ggplot2::aes(color = model),
        columns=3:6,
        lower = list(continuous = wrap("points", alpha = 0.5)),
        diag = list(continuous = wrap("densityDiag", alpha=0.5)),
        upper = list(continuous = "blank"))+
    theme(panel.margin=grid::unit(0,"lines"))
## http://stackoverflow.com/questions/14711550/is-there-a-way-to-change-the-color-palette-for-ggallyggpairs-using-ggplot
cbscales <- list(scale_colour_discrete(),
                 scale_fill_discrete())
for (row in seq_len(ggp1$nrow))
    for (col in seq_len(ggp1$ncol))
        ggp1[row, col] <- ggp1[row, col] + cbscales[[(row==col)+1]]
ggp1
```

```{r corrsum,message=FALSE,cache=TRUE}
m1 <- rename(melt(sum_list[["sum_mat"]],id.vars=c("model","run")),
             sumvar=variable,sumval=value)
m2 <- rename(melt(ltab,id.vars="run"),
             LHSvar=variable,LHSval=value)
mL2 <- full_join(m1,m2)
```

## Sensitivity

```{r corrsum_plot,dev="png",echo=FALSE,fig.width=10,fig.height=6}
ggplot(mL2,aes(LHSval,sumval,colour=model))+
    geom_point(pch=".",alpha=0.5)+
    facet_grid(sumvar~LHSvar,scales="free")+
    geom_smooth(se=FALSE)+labs(x="",y="")
```

## Conclusions/questions: details

- essentially:
implicit $\approx$ pairform < instswitch < instswitch+epc $\approx$ pairform+epc
- why no low-eq-vir tail for pairform+epc?
- why is pairform, not instswitch, closest to implicit?

## Conclusions/questions: big picture

- how do we decide what to pick on?
- how does scaling $r$ help/hurt?

## References {.refs .columns-2}