### 2016-03-12 ms/ms_parms.tex

\begin{table}[ht]
\caption{Parameter ranges/values}
\centering
\begin{tabular}{c p{2in} c l}
\hline 
Notation & Description & Range/Value & Source\\
\hline % inserts single horizontal line
$\rho$ & Partnership formation rate & 1/10-2/5 & \cite{champredon_hiv_2013} \\
$c$ & Partnership dissolution rate & 1/15-1/5 & \cite{champredon_hiv_2013} \\
$c_u/c_w$ & Relative contact rate uncoupled & 1/5-5 & Assumption \\
$c_e/c_w$ & Relative contact rate extra-couple & 0.01-1 & \cite{champredon_hiv_2013} \\
$\beta_P$ & Rate of transmission during primary infection & 1.31-5.09 & \cite{hollingsworth_hiv1_2008} \\
$\beta_D$ & Rate of transmission during high transmission disease stage & 0.413-1.28 & \cite{hollingsworth_hiv1_2008} \\
$D_P$ & Duration of primary infection & 1.23/12-6/12 & \cite{hollingsworth_hiv1_2008} \\
$D_D$ & Duration of high transmission disease stage & 4.81/12-14/12 & \cite{hollingsworth_hiv1_2008} \\
$\beta_{max}$ & Maximum rate of transmission during asymptomatic stage & 0.317 & \cite{shirreff_transmission_2011} \\
$\beta_{50}$ & SPVL at which infectiousness is half maximum & 13938 & \cite{shirreff_transmission_2011} \\
$\beta_k$ & Hill coefficient: steepness of increase in infectiousness as a function of SPVL & 1.02 & \cite{shirreff_transmission_2011} \\
$D_{max}$ & Duration of primary infection & 25.4 & \cite{shirreff_transmission_2011} \\
$D_{50}$ & SPVL at which duration of asymptomatic infection is half maximum & 3058 & \cite{shirreff_transmission_2011} \\
$D_{k}$ & Hill coefficient: steepness of decrease in duration as a function of SPVL & 0.41 & \cite{shirreff_transmission_2011} \\
$\sigma_M$ & Mutation standard deviation of $\log_{10}$ SPVL & 0.12 & \cite{shirreff_transmission_2011} \\
$\alpha_{min}$ & Minimum $\log_{10}$ SPVL & 2 & \cite{shirreff_transmission_2011}\\
$\alpha_{max}$ & Maximum $\log_{10}$ SPVL & 2 & \cite{shirreff_transmission_2011}\\
$n$ & Number of strains & 21 & Assumption\\
\hline
\end{tabular}
\label{table:parmsTable}
\end{table}