### 2016-05-10 ms/ms.Rnw
considerable work on ms

\documentclass{article}

%\bibpunct{(}{)}{;}{a}{}{,}
\usepackage{natbib}
\usepackage{palatino}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage[usenames,dvipsnames]{xcolor} %% for non-standard cite color
\usepackage{graphicx}
\usepackage[colorlinks = true, linkcolor = blue, urlcolor=blue,citecolor = MidnightBlue]{hyperref}
\usepackage{url}
\usepackage{enumerate}
%\usepackage[normalem]{ulem}
\usepackage[left=50pt,
            textwidth=345pt,
            marginparsep=25pt,
            marginparwidth=124pt]{geometry}
%\usepackage[margin=1in]{geometry}

\newcommand{\etal}{et al.}
\newcommand{\shir}{Shirreff et al.}%{\citeauthor{shirreff_transmission_2011}}
\newcommand{\champr}{Champredon et al.}%{\citeauthor{champredon_hiv_2013}}
\newcommand{\tsub}[2]{#1_{{\textrm{\tiny #2}}}}
\newcommand{\khalf}{\left(\frac{1}{2}\right)^{\delta_{ij}}}  % (1/2)^kronecker
\newcommand{\Lspvl}{$\log_{10}$ SPVL}
\newcommand{\rzero}{{\cal R}_0}
\usepackage{changes}
\usepackage[backgroundcolor=lightgray,textsize=small]{todonotes}

<<setup,echo=FALSE>>=
library("knitr")
opts_chunk$set(echo=FALSE,error=FALSE,fig.width=8,fig.height=4) #
## https://github.com/yihui/tikzDevice/issues/46 
local({
    hook_document <- knitr::knit_hooks$get('document')
    knitr::knit_hooks$set(document = function(x){
        res <- hook_document(x)
        gsub('\\usepackage[]{color}', '\\usepackage[usenames,dvipsnames]{xcolor}', res, fixed = TRUE)
  })
})
@

<<pkgs,warning=FALSE,message=FALSE>>=
library("ggplot2"); theme_set(theme_bw())
zero_margin <- theme(panel.margin=grid::unit(0,"lines"))
library("MASS")
## use Dark2 rather than Set1 because colour #6 of Set1 is yellow (ugh/too light)
scale_colour_discrete <- function(...,palette="Dark2") scale_colour_brewer(...,palette=palette)
scale_fill_discrete <- function(...,palette="Dark2") scale_fill_brewer(...,palette=palette)
library("gridExtra")
library("plyr")  ## for ldply()
library("dplyr") ## for full_join()
library("reshape2")  ## for melt()
library("GGally")
library("knitr")
## devtools::install_github("lionel-/ggstance")
library("ggstance")
@

<<loadstuff,cache=TRUE>>=
load("../simdata/combine_ev_LHS4.rda")
@

\author{Sang Woo Park and Ben Bolker}
\title{Effects of epidemiological structure on the transient evolution of HIV virulence}

\begin{document}
\maketitle

\abstract{The evolutionary dynamics of parasite virulence over the
  course of an emerging epidemic have important implications both for
  our basic understanding of epidemiological dynamics and,
  potentially, for the outcomes of public health interventions. In
  general changes in fitness landscapes over the course of an epidemic
  will select for higher virulence during the early,
  exponential-growth phase of the epidemic, but quantitative outcomes
  can depend sensitively on biological details and the structure of
  mathematical models used to capture them.  Fraser, Shirreff, and
  co-workers have proposed a series of models for eco-evolutionary
  dynamics of HIV that are relatively detailed in their portrayal of
  the tradeoffs between transmission and virulence (mediated by
  set-point viral load, SPVL) and their heritability between
  hosts. However, these models use very simple implicit
  representations of the transmission process that ignore the
  partnership dynamics that previous research has found to be critical
  in predicting epidemics of sexually transmitted diseases.  We
  explore models that combine HIV virulence tradeoffs with a range of
  epidemiological structures, modeling partnership formation and
  dissolution and allowing for individuals to transmit disease outside
  of partnerships. We assess summary statistics such as the peak value
  of virulence (SPVL) and the time at which the peak occurs across all
  models and across a Latin hypercube sample that captures a realistic
  range of partnership dynamic parameters for sub-Saharan Africa. In
  order to account for the different interpretations of parameters
  across model structures, we scale all parameter sets to constrain
  the simulated epidemic growth rate to be identical, matching a
  realistic baseline value. Our primary result is that, for this
  particular model setting, the simplest random-mixing structure is
  actually the best approximation to the most realistic model; this
  surprising outcome occurs because the dominance of extra-pair
  contact in the realistic model tends to mask the effects of
  partnership structure.
}
\section{Introduction}

The evolution of pathogen virulence is a fundamental process in
evolutionary biology, of both theoretical and (potentially) practical
importance. The trade-off theory \citep{Ebert1999} --- which
postulates that parasite virulence can be explained as the long-term
evolutionary outcome of a saturating relationship between parasite
clearance rate and transmission rate --- has been criticized
\citep{EbertBull2003,alizon_adaptive_2015}, but has also been
successfully applied in a variety of host-pathogen systems \citep{Dwyer+1990,mackinnon1999genetic,jensen2006empirical,deroode2008virulence}. One
particularly interesting application of these ideas is the work by
Fraser \etal\ showing that HIV appears to satisfy the prerequisites of
the tradeoff theory: in a study of discordant couples (i.e. long-term
sexual partnerships with one infected and one uninfected partner), HIV
virulence as measured by the rate of progression to AIDS was both
heritable and covaried with the set-point viral load (i.e., the
characteristic virus load measured in blood during the intermediate
stage of infection), which in turn predicted the probability of
within-couple transmission
\citep{Fraser+2007,fraser_virulence_2014}\todo{read/check fraser+
  2014!; check vs. \cite{herbeck_is_2012}}. Subsequent studies
\citep{shirreff_transmission_2011,herbeck_hiv_2014} used these data to
parameterize mechanistic models of HIV virulence evolution, suggesting
that HIV invading a novel population would initially evolve increased
virulence, peaking after approximately [XXX] years and then declining
slightly to a long-stable virulence level.

\todo{\cite{gras2009viral} and \cite{muller2009increasing} say that mean virulence has been increasing in Netherlands and Italy. Empirical support? Worth checking out maybe? I think these are similar to what \cite{herbeck_is_2012} says about the evolutionary trend but more specific...}

The work of \shir, and particularly the predicted
transient peak in HIV virulence midway through the epidemic,
highlights the importance of interactions between epidemiological and
evolutionary factors \citep{day_virulence_2004,alizon_price_2009}.
However, despite the attention to mechanistic detail at the individual
or physiological level, the epidemiological structures used in these
models are relatively simple. \todo{don't forget to check other recent papers: esp \cite{payne_impact_2014}}

\todo{\cite{lythgoe2013hiv} --- multi-strain model + within host dynamics but assumes random mixing. This one considers multi-strains existing within a host. Another complication that is important but this paper also ignores the partnership dynamics...}

\todo{\cite{van2014immuno} --- seems even more complicated but doesn't seem like it mentions anything about partnership dynamics... HIV phases are based on Fraser and Hollingsworth et al}

\todo{\cite{herbeck2016evolution} --- seems better than the previous herbeck et al paper but still weird and is not published (bioRxiv)..} As we discuss in detail below, the
existing models of HIV eco-evolutionary dynamics either use implicit
models that incorporate the average effects of within-couple sexual
contact --- without representing the explicit dynamics of pair
formation and dissolution or accounting for extra-partnership contact
--- or use an agent-based formulation with parameters that effectively
lead to random mixing among infected and uninfected individuals. Here
we explore the effects of incorporating \emph{explicit}
epidemiological structure in eco-evolutionary models.

We add complexity to the epidemiological model following the general approach of \cite{champredon_hiv_2013}; individuals join and leave partnerships at a specified rate, and can have sexual contact both within and outside of established partnerships. At the same time, our analysis somewhat simplifies the models of \shir, for computational tractability; we check that our qualitative results are not sensitive to these simplifications. In order to explore how virulence evolution depends on epidemiological structure, we consider a series of models with increasing levels of complexity. In order to avoid dependence of the results on a particular set of parameters --- as we explain below, finding matching sets of parameters across models with widely differing epidemiological structures is challenging --- we evaluate our models across a wide range of parameters, again following \champr\ in using a Latin hypercube design. For each model run, we compute a set of metrics (peak virulence, timing of virulence peak, equilibrium virulence) that summarize the evolutionary trajectory of a simulated HIV epidemic.

\section{Methods}

\todo{more papers on SPVL heritability (\emph{maybe} include): \cite{alizon_phylogenetic_2010,leventhal_potential_2016}}

As our primary goal is to explore how different epidemiological structures (i.e. partnership dynamics and contact structures) affect our conclusions about the evolution of virulence, our models use a simplified description of within-host dynamics and heritability derived from 
\shir's multi-strain evolutionary model. Like \shir, we use a simple susceptible-infected-susceptible demographic formulation; rather than modeling birth and death (or more specifically, recruitment into the sexually active population and death), we assume that whenever an individual dies from infection, another enters the susceptible compartment.

\subsubsection{Infection dynamics}

Like \shir, we focus on the evolution of mean $\log_{10}$ set-point viral load, SPVL (which we denote as $\alpha$), rather than virulence (i.e. rate of progression to AIDS) itself. 
In contrast to \shir, we use a single-stage disease model instead of accounting explicitly for progression through the three main stages of HIV infection (primary, asymptomatic, and disease), and we use a simple exponentially distributed infectious period instead of a more realistic Weibull-distributed infectious period. We account for varying transmission rates and durations of each disease stage by summing the durations of three stages (again based on \shir's model) and taking the duration-weighted average of transmission rates of three stages. Thus the within-couple transmission rate, $\beta$, for our models is given by:
\begin{equation}
\beta (\alpha) = \frac{D_P \beta_P + D_A (\alpha) \beta_A (\alpha) + D_D \beta_D}{D_P + D_A (\alpha) + D_D}
\end{equation}
where the duration of infection ($D_P$ and $D_D$) and rate of transmission ($\beta_P$ and $\beta_D$) of the \textbf{P}rimary and \textbf{D}isease stages
of infection are independent of the host's SPVL. Following \shir, the duration of infection ($D_A$) and rate of transmission ($\beta_A$) for the \textbf{A}symptomatic stage are Hill functions of the SPVL:

\begin{equation}
\begin{split}
D_A(\alpha) &= \frac{\tsub{D}{max} D_{50} ^{D_k}}{V_\alpha ^{D_k} + D_{50}^{D_k}}, \\
\beta_A(\alpha) &= \frac{\tsub{\beta}{max} V_\alpha ^ {\beta_k}}%
{V_\alpha^{\beta_k} + \beta_{50} ^{\beta_k}},
\end{split}
\end{equation}
where $V_{\alpha} = 10^\alpha$. 
The \textbf{u}ncoupled and \textbf{e}xtra-couple transmission rates are scaled by multiplying the \textbf{w}ithin-couple transmission rate $\beta$ by the contact ratios $c_u/c_w$ and $c_e/c_w$.

\subsubsection{Mutation}

Like \shir\ we incorporate a between-host mutation process in the SPVL, but simplify \shir's evolutionary model slightly by using a one-to-one genotype-phenotype mapping. 
The mutational process in our model is directly taken from \shir. Over the course of infection, mutation occurs within the host. However, it is assumed that SPVL of an infected individual is determined by the SPVL at the time of infection for simplicity (and is not further affected by within-host mutation). Instead, the mutational effect takes place when an infected individual transmits the virus to a susceptible individual. First, the distribution of \Lspvl\ is discretized into a vector:

\begin{equation}
\alpha_i =  (\tsub{\alpha}{max} - \tsub{\alpha}{min})\frac{(i-1)}{n-1} + \tsub{\alpha}{min} \qquad i = 1,2,3, \dots n.
\end{equation}
Then, we construct an $n$ by $n$ mutational matrix, $M$ --- which is multiplied with the transmission term ---  so that $M_{ij}$ is the probability that a newly infected individual will have \Lspvl\ of $\alpha_j$ given that the infector has \Lspvl\ of $\alpha_i$. Finally, the probabilities are normalized so that each row sums to 1:

\begin{equation}
M_{ij} = \frac{\Phi(\alpha_j + d/2;i) - \Phi(\alpha_j - d/2;i)}{\Phi(\tsub{\alpha}{max} + d/2;i) - \Phi(\tsub{\alpha}{min} - d/2;i)},
\end{equation}

where $\Phi(x;i)$ is the Gaussian cumulative distribution function with mean $\alpha_i$ and variance of $\sigma_M^2$, and $d = (\tsub{\alpha}{max} - \tsub{\alpha}{min})/(n-1)$. Transmission rate and disease induced mortality rates are discretized into a vector as well:

\begin{equation}
\begin{aligned}
\beta_i &= \beta(\alpha_i)\\
\lambda_i &= \frac{1}{D_P + D_A (\alpha_i) + D_D}
\end{aligned}
\end{equation}

\subsubsection{Contact structure and partnership dynamics}

We developed six multi-strain evolutionary models, designed to cover a gamut between \champr's relatively realistic and \shir's relatively simplistic epidemiological structures, each of which is based on different assumptions regarding contact structure and partnership dynamics. Specifically, we focus on the effects of the assumptions of (1) instantaneous vs. non-instantaneous partnership formation and (2) zero vs. positive extra-partnership sexual contact and transmission on the evolution of mean \Lspvl.

Our first four models consider explicit partnership dynamics and are based on \champr's model. Models 1 and 2 assume non-instantaneous partnership formation (i.e. individuals spend some time uncoupled, outside of partnerships) and consist of five states that are classified by infection status and partnership status. $S$ is the number of single (uncoupled) susceptible individuals, and $I$ is the number of single infected individuals. $SS$ is the number of susceptible-susceptible couples, $SI$ is the number of serodiscordant (susceptible-infected) couples, and $II$ is the number of concordant positive (infected-infected) couples. Model 1 includes extra-partnership contact (with both uncoupled individuals and individuals in other partnerships) whereas model 2 only considers within-couple transmission. Models 3 and 4 assume instantaneous partnership formation and thus consist of only the three partnered states: $SS$, $SI$, and $II$. Parallel to model 1 and 2, model 3 includes extra-partnership contact (now only with individuals in other partnerships, since uncoupled individuals don't exist in this model) and model 4 only considers within-couple transmission.

In contrast, models 5 and 6 are not explicitly structured.  Model 5 is an implicit serial monogamy model based on the epidemiological model used by \shir. It is actually a random mixing model that consist of only two states, $S$ and $I$, and does not consider explicit partnership dynamics. However, to simulate the effect of instantaneous partnership formation, it uses an adjusted transmission rate that is derived from approximated basic reproduction number of a serial monogamy model \citep{hollingsworth_hiv1_2008}. Finally, model 6 is a simple random-mixing model.

The base model (i.e. model 1) for the first four models is an extension of \champr's model. Individuals in single compartment acquire a partner at a rate, $\rho$, and partnerships dissolve at a rate, $c$. Infected individuals in a discordant partnership infect susceptible partner at a rate $\beta$ (within-couple transmission rate) and susceptible individuals outside the partnership at a rate $c_e$ (extra-couple transmission rate). Likewise, a single infected individual can infect any susceptible individuals at a rate $c_u$ through uncoupled mixing. Extra-couple and uncoupled transmission are modeled in a same way as \champr's model. All the details have been adapted to a multi-strain scenario. Model 2, 3, and 4 are derived from the base model by removing epidemiological details (partnership formation and uncoupled/extra-couple contact). Model details are explained in the appendix.

\input{ms_parms.tex}

\subsection{Latin hypercube sampling}

Despite considerable effort by many researchers \citep[e.g.][]{hollingsworth_hiv1_2008,champredon_hiv_2013}, the parameters determining the rate and structure of sexual partnership change and contact are still very uncertain; this led \champr\ to adopt a Latin hypercube sampling (LHS) strategy \citep{blower_drugs_1991} that evaluates model outcomes over a range of parameter values. In order to make sure that our comparisons among models apply across the entire space of reasonable parameter values, and in order to evaluate the differential sensitivity of different model structures to parameter values, we follow a similar protocol and perform LHS over a parameter set including both the transmission and duration parameters ($\beta_P$, $D_P$, $\beta_D$, $D_D$) and contact/partnership parameters ($\rho$, $c$, $c_u/c_w$, and $c_e/c_w$). We do not allow for uncertainties in parameters that are directly related to the evolutionary process ($\tsub{\beta}{max}$, $\beta_{50}$, $\beta_k$, $\tsub{D}{max}$, $D_{50}$, $D_k$, $\sigma_M$), using \shir's point estimates throughout.

Latin hypercube sampling is done as in \champr. For each parameter, $z$, its range is divided into $N = 1000$ equal intervals on a log scale:

\begin{equation}
z_i = \exp\left(\log(\tsub{z}{min}) + [\log(\tsub{z}{max}) - \log(\tsub{z}{min})] \frac{i-1}{N-1}\right) \qquad i = 1,...,N.
\end{equation}
For simplicity, we assume that all parameters are uniformly distributed on the log scale. Following the vectorization of a parameter range, a matrix is constructed so that each column contains a vector of a parameter series which it represents ($z_1$,...$z_N$). Then, each column is replaced with a random permutation series of itself. Now, each row contains a different parameter set that is used for each simulation run.

Table~\ref{table:parmsTable} gives the ranges of the model parameters used for LHS. Parameter ranges regarding contact and partnership dynamics ($\rho$, $c$, and $c_e/c_w$) are taken from \champr, whereas those regarding infection ($\beta_P$, $D_P$, $\beta_D$, and $D_D$) are taken from \cite{hollingsworth_hiv1_2008}. The remaining parameters are taken from \shir.

The one completely new parameter in our model, the ratio of uncoupled to within-couple transmission $c_u/c_w$, is needed to more flexibly contrast uncoupled and extra-couple transmission dynamics within multi-strain models (Appendix); we need to pick a reasonable range for it. \champr\ assume that the effective within-couple contact rate and effective uncoupled contact rate have the same range of 0.05 - 0.25.  Given \champr's parameter range, the possible maximum and minimum values of $c_u/c_w$ are 5 and 1/5. Therefore, we use 1/5-5 as the range for the parameter $c_u/c_w$. Although this adds more uncertainty to the parameter $c_u$ --- \champr's range implies a 5-fold difference whereas ours gives a 25-fold difference --- as there is not much known about the uncoupled transmission rate, we consider the wider range appropriate.

\subsection{Simulation runs}

One of the most difficult parts of model comparison is finding parameter sets that are commensurate against many different model structures. For the most part, our models are too complex to easily derive analytical correspondences among them. Given a numerical criterion, such as $r$ (initial exponential growth rate) or $\rzero$ (intrinsic reproductive number), we can adjust one or more parameters by brute force to ensure that all of the models match according to that criterion. While $\rzero$ is often considered the most fundamental property of an epidemic, and might thus seem to be a natural goal, here we focus on matching the initial growth rate $r$ for several reasons. First, our primary interest is in the transient evolutionary dynamics of virulence, which are more strongly affected by $r$ than $\rzero$ \todo{CITE?}. Second, $r$ is in general more directly observable in real epidemics; $r$ can be estimated by simply fitting an exponential curve to the initial incidence or prevalence curves \citep{ma_estimating_2014}, while $\rzero$ typically requires either (1) knowledge of \emph{all} epidemic parameters or (2) relatively sophisticated back-calculation based on $r$ and knowledge of the serial interval or generation interval of the disease \citep{wallinga_how_2007}. Thus, we scale a parameter so that every run has the same initial exponential growth rate in the incidence.

In order to allow for all models to have equal initial exponential growth rate, $r$, we need to pick a parameter, $s$, so that $\lim_{s\to 0} r(s) = 0$ and $\lim_{s\to\infty}  r(s) = \infty$. As adjusting either partnership change rate (i.e. partnership formation and dissolution rate) or transmission rate does not fulfill this requirement for certain models, we decided to scale partnership change rate and dissolution rate by the same factor of $\gamma$: $\tsub{\beta}{adj} =  \gamma \tsub{\beta}{base}$, $\tsub{c}{adj} = \gamma \tsub{c}{base}$, $\tsub{\rho}{adj} = \gamma \tsub{\rho}{base}$. Since transmission rate is adjusted by the scale of $\gamma$, uncoupled and extra-couple transmission rates are adjusted as well. For models 3, 4, and 5, all of which assume instantaneous partnership (and hence do not track single individuals), only the transmission rate and partnership dissolution rate (in this case equivalent to the partnership change rate) are adjusted.

We run each model for each of 1000 parameter sets chosen by Latin hypercube sampling \todo{CITE Blower et al - or at first mention of LHS}, with fixed starting conditions
of mean \Lspvl\ of 3 and epidemic size of $10^{-4}$. After each run, initial exponential growth rate is calculated. Then, parameters are scaled so that the initial exponential growth rate is scaled to 0.04, which is approximately equal to that of Shirreff et al's model.

For each model we derive the following summary statistics: peak virulence, peak time, equilibrium virulence, and relative peak virulence. The transient phase of an epidemic is often characterized by high virulence, and we define peak virulence as the maximum virulence during this phase. It is simply calculated by taking the maximum value from the virulence trajectory, and peak time is the time at which the maximum value is reached. Once the epidemic enters the endemic phase, evolution of virulence stabilizes and reaches equilibrium. Equilibrium virulence is calculated by taking the mean virulence at 4000 years. Although most simulations reach equilibrium much earlier, we set our time horizon at a much later date as some simulation runs have slow rate of evolution depending on the parameter set and model assumptions.

We focus on these statistics for the following two reasons. First of all, knowing the possible ranges for the peak virulence allows us to estimate the worst-case scenario for the HIV and other sexually transmitted disease epidemics. Pathogens may already have evolved towards high virulence during the early stages of an epidemic, by the time it is observed by public health authorities. Understanding how virulent a pathogen can evolve before an epidemic begins can be helpful for controlling the disease. Furthermore, knowing the initial virulence, peak virulence, timing of the peak virulence, and equilibrium virulence provide sufficient detail to identify the shape of the virulence trajectory. During an epidemic outbreak, it is difficult to observe virulence evolution. Specifically, in the case of HIV and other sexually transmitted diseases, slow evolutionary time-scale makes observing changes in the mean virulence even more challenging. Knowing the ranges of these statistics can help real-time virulence evolution prediction during an epidemic less troublesome.\todo{needs a bit of work}

\section{Results}

Our simplifications of \shir's model reproduce its qualitative behaviour --- in particular, its predictions of virulence dynamics --- reasonably well. As $r$ decreases from 0.084 to 0.42 (the latter value matching the initial rate of increase in prevalence in \shir's full model) the initial trajectory of increasing virulence brackets the rate from the original model (Figure~\ref{fig:panel3}a). However, our model produces lower peak virulence ($\approx 4.3$ vs. $\approx 4.6$ \Lspvl) 
and equilibrium virulence $\approx 4.25$ vs. $\approx 4.5$ \Lspvl) than Shirreff's, even for matching initial incidence trajectories (i.e., $r=0.042~\textrm{year}^{-1}$).

Changing the initial infectious density ($I(0)$), while it produces the expected changes in the initial epidemic trajectory (Supplementary material), has little effect on the virulence trajectory, making the virulence peaks slightly later and larger as $I(0)$ decreases. Decreasing $I(0)$ allows a longer epidemic phase before the transition to endemic dynamics (Figure~\ref{fig:panel3}b). Decreasing the initial virulence similarly but more strongly leads to progressively later, larger peaks in virulence (Figure~\ref{fig:panel3}c).

<<setup2,echo=FALSE>>=
orig_sum_labs <- c("peak_time","peak_vir","eq_vir","rel_peak")
new_sum_labs <- c("peak time","peak virulence","equil virulence","relative peak")
fixfac2 <- function(x) factor(x,
                       levels=orig_sum_labs,
                       labels=new_sum_labs)
@

<<panel3,fig.cap="Baseline dynamics. Time series of mean population virulence ($\\log_{10}$ SPVL). (a) Shirreff model, effects of varying $r$. (b) Effects of varying initial infectious density $I(0)$. (c) Effects of varying initial mean virulence $\\alpha(0)$. Parameters",dev="tikz",fig.height=3,fig.width=8,cache=TRUE>>=
for (i in c("","2","3")) 
    load(sprintf("../simdata/ev_LHS_resS%s.rda",i))
Ls <- load("../simdata/shirreff_data.rda")
tvec <- seq(from=1,by = 0.1,length.out=nrow(vir_matS))
dimnames(vir_matS) <- list(Time=tvec,
                           run=paste0("$r=",c(0.021,0.042,0.084),"$"))

d_littleR_m <- melt(vir_matS,value.name="Mean.VL")
shirreff_df_m <- subset(res_sum_df,run==1,select=c(run,Time, Mean.VL))
shirreff_df_m$run <- "Shirreff"
df_tot <- rbind(shirreff_df_m,d_littleR_m)
vlab <- "mean virulence ($\\log_{10}$ SPVL)"
tlab <- "time (years)"
g1 <- ggplot(df_tot,aes(x=Time,y=Mean.VL,colour=run))+
    geom_line()+labs(x=tlab,y=vlab)+
    theme(legend.position=c(0.8,0.2))
###
dimnames(vir_matS2) <- list(Time=tvec,
               run=c("$10^{-3}$", "$10^{-4}$", "$10^{-5}$"))
g2 <- g1 %+% melt(vir_matS2,id.var="Time",value.name="Mean.VL")
###
dimnames(vir_matS3) <- list(Time=tvec,
                            run=c("$2.5$", "$3$", "$3.5$"))
g3 <- g1 %+% melt(vir_matS3,id.var="Time",value.name="Mean.VL")
limits <- c(2.5,4.7)
breaks <- seq(limits[1],limits[2], by = 0.5)
afun <- function(label,size=8) {
    annotate(geom="text",label=label,x=-Inf,y=Inf,
             size=size,vjust=0.98,hjust=0.02)
}
g1.y <- g1 + scale_y_continuous(limits = limits, breaks = breaks)+
    scale_colour_discrete(name="run")+afun("a")
g2.y <- g2 + scale_y_continuous(limits = limits, breaks = breaks)+
    scale_colour_discrete(name="$I(0)$")+afun("b")
g3.y <- g3 + scale_y_continuous(limits = limits, breaks = breaks)+
        scale_colour_discrete(name="$\\alpha(0)$")+afun("c")
grid.arrange(g1.y, g2.y, g3.y, nrow=1)
@

Across the entire range of parameters covered by the LHS analysis, all of the classes of models we considered produce qualitatively similar virulence trajectories (Figure~\ref{fig:virtraj}). Although the speed of virulence evolution varies, leading to wide variation in the peak virulence (means ranging from approximately 3.75 to 4.5 \Lspvl), virulence peaks in all models between 200 and 300 years.

<<virtraj,fig.cap="Envelopes of virulence trajectories under all models. All models were run until $t=4000~\\textrm{years}$; truncated series are shown here.",fig.width=6,cache=TRUE,message=FALSE,warning=FALSE,dev="tikz">>=
ss <- subset(sum_list[["vir_mat"]],tvec<750)
gg_virtraj <- ggplot(ss,
                     aes(tvec,mean,ymin=lwr,ymax=upr))+
    geom_line(aes(colour=model,linetype=model),lwd=1.5)+
    geom_ribbon(aes(fill=model),alpha=0.3)+
    labs(x="time (years)",y=vlab)
## scale_x_continuous(limits=c(0,2500))
print(gg_virtraj)
@

<<unidist,fig.cap="Univariate distributions of summary statistics. The distribution of equilibrium virulence for the random mixing model is very narrow, and has been replaced by a point in order to preserve the vertical axis scaling.",fig.width=8,fig.height=4.8>>=
sL <- transform(sum_list[["sum_mat"]],rel_peak=peak_vir/eq_vir)
mL <- melt(sL,id.vars=c("model","run"))
## horrible hack (but doesn't help); subsample
## w <- which(mL$model=="random")
## mLw <- mL[-sample(w,size=length(w)*9/10,replace=FALSE),]
w <- with(mL,which(model=="random" & variable=="eq_vir"))
rval <- mean(mL$value[w])
mLw <- droplevels(mL[-w,])
mLw$variable <- fixfac2(mLw$variable)
gg_univ <- ggplot(mLw,aes(value,model,fill=model))+
    geom_violinh(width=1)+
    facet_wrap(~variable,scale="free_x")+
    guides(fill=guide_legend(reverse=TRUE))+
    labs(y="")+
    geom_point(data=data.frame(model="random",variable="equil virulence",
                               value=rval),pch=22,size=3,show.legend=FALSE)
print(gg_univ)
@

<<pairplot_funs,echo=FALSE>>=
trim_gg <- function(gg) {
    n <- gg$nrow
    gg$nrow <- gg$ncol <- n-1
    v <- 1:n^2
    gg$plots <- gg$plots[v>n & v%%n!=0]
    gg$xAxisLabels <- gg$xAxisLabels[-n]
    gg$yAxisLabels <- gg$yAxisLabels[-1]
    return(gg)
}

tweak_legends_gg <- function(gg,pos=1.5) {
    n <- gg$nrow
    for (i in 1:n) {
        for (j in 1:n) {
            inner <- getPlot(gg,i,j)
            if (i==1 & j==1) {
                inner <- inner + ggplot2::theme(legend.position=c(pos,0.5)) +
                    guides(colour = guide_legend(override.aes = list(size=6)))
            } else {
                inner <- inner + ggplot2::theme(legend.position="none")
            }
            gg <- putPlot(gg,inner,i,j)
        }
    }
    return(gg)
}
tweak_colours_gg <- function(gg) {
    n <- gg$nrow
    for (i in 1:n) {
        for (j in 1:n) {
            inner <- getPlot(gg,i,j)
            inner <- inner+scale_colour_brewer(palette = 'Dark2')
            gg <- putPlot(gg,inner,i,j)
        }
    }
    return(gg)
}
@ 
<<pairplot,echo=FALSE,fig.cap="Pairs plot: bivariate relationships among summary statistics for each model structure. Dashed line in equilibrium vs. peak virulence plot shows 1:1 line.",fig.width=7,fig.height=7>>=
sL2 <- sL
names(sL2)[na.omit(match(orig_sum_labs,names(sL)))] <- gsub(" ","_",new_sum_labs)
    ## paste0("`",new_sum_labs,"`")
ggp1 <- ggpairs(sL2,
        mapping = ggplot2::aes(color = model),
        columns=3:6,
        legends=TRUE,
        lower = list(continuous = wrap("points",size=0.1)), ## alpha = 0.3,size=0.5)),
        diag = list(continuous = "blankDiag"),
        upper = list(continuous = "blank"))
## http://stackoverflow.com/questions/14711550/is-there-a-way-to-change-the-color-palette-for-ggallyggpairs-using-ggplot
## have to change plot one panel at a time
ggp2 <- tweak_colours_gg(tweak_legends_gg(trim_gg(ggp1)))
## add 1-to-1 line
inner <- getPlot(ggp2,2,1)
inner <- inner+geom_abline(intercept=0,slope=1,lty=2)
ggp2 <- putPlot(ggp2,inner,2,1)
## tweak scale
inner <- getPlot(ggp2,3,1)
inner <- inner+scale_x_continuous(breaks=seq(2.5,4.5,by=0.5))
ggp2 <- putPlot(ggp2,inner,3,1)
print(ggp2)
@

Our chosen summary statistics (peak time, peak virulence, equilibrium
virulence, and relative peak virulence) all vary considerably across models
\ref{fig:unidist}.
We first consider the models of intermediate realism (implicit,
instantaneous-switching with and without extra-pair contact, and
pair formation without extra-pair contact). Some parameter
sets for these models lead to low equilibrium virulence ($\approx 2.5 \log_{10}$ SPVL)
these same sets lead to correspondingly low
peak virulence ($< 3.5 \log_{10}$ SPVL) and early peak times (before 200 years), 
but high relative peaks ($>1.3$)
(Figure~\ref{fig:pairplot}, leftmost column) because the equilibrium virulence is low.
At the opposite extreme, parameter sets that produce high equilibrium virulence 
also produce late peaks ($> 200 \text{ years}$), 
high peak virulence, and low relative peaks ($\approx 1.05$).
The pair-formation without extra-pair contact and implicit models
occasionally have parameter sets that select for such low virulence across
the board that they never exceed their initial virulence, leading to a tail
of peak times near zero.

% \end{document}

The most striking aspect of the univariate comparisons in
Figure~\ref{fig:unidist}, (and the bivariate comparisons in
Figure~\ref{fig:pairplot}) is the match between the output of the
least (random mixing) and the most complex (pair formation with
extra-pair contact) models. The random-mixing model has lower
variability, because it is unaffected by uncertainty in pair formation
and extra-pair contact parameters, but otherwise the virulence
dynamics of these two extreme models are remarkably similar.
This phenomenon is driven by the strength of extra-pair contact in the
pairform+epc model. When individuals spend time uncoupled between
partnerships, and when these single individuals can transmit disease
to coupled individuals, the resulting unstructured mixing overwhelms
the effect of structured mixing within couples, leading to mixing
that is effectively close to random.

Plotting the bivariate result distributions (Figure~\ref{fig:pairplot})
shows that most of the relationships among the summary statistics
are monotonic, except those involving the relative peak virulence
(bottom row). Changes in parameters that increase the equilibrium
virulence initially increase the peak virulence even more, so
that the relative peak virulence increases as well, but beyond
an equilibrium virulence of about 2.5 \Lspvl\ the peak virulence
increases slower than the equilibrium virulence, leading to a
decrease in the relative peak virulence.

The bivariate relationships also help distinguish the results of 
different models with similar univariate distributions. While the
relationship between equilibrium virulence and peak time is
similar for all model structures (top left panel), the other
relationships are more separate. In particular, the implicit
and pair-formation (without extra-pair contact) are very similar
to each other, but distinct from the other models. We still do
not have a convincing explanation for this distinction; we
would have expected the implicit model to be most similar to the
the instantaneous-switching model without extra-pair contact,
which most closely matches its derivation. However, we note
that the implicit model derivation is based on defining
the force of infection to match a scaled version of $\rzero$,
and as such would be expected to match the equilibrium behaviour
but not necessarily the epidemic-phase behaviour of a model
with explicit partnership dynamics.

<<testrev,echo=FALSE,eval=FALSE>>=
set.seed(101)
zz <- data.frame(x=rnorm(3000),f=factor(rep(letters[1:3],each=1000)))
g0 <-  ggplot(zz,aes(x=f,y=x,fill=f))+geom_violin()
grid.arrange(
    g0,
    g0+ guides(fill=guide_legend(reverse=TRUE)),
    nrow=2)
@

<<plot_sens,dev="png",echo=FALSE,out.width="\\textwidth",fig.width=8,fig.height=3.5,dpi=240,fig.cap="Sensitivity plot. For each parameter in the Latin hypercube sample and each summary statistic, shows the distribution (points) and trend (smooth line) of the summary statistic as a function of the \\emph{unscaled} parameter value, i.e. prior to adjusting the parameters to achieve the standard initial epidemic growth rate.">>=
mL3 <- subset(mL2,
       !((model=="implicit" &
          LHSvar %in% c("rho_base","c_u_ratio","c_e_ratio")) |
         (model=="random" &
          LHSvar %in% c("rho_base","c_u_ratio","c_e_ratio","c_mean_base")) |
         (model=="instswitch" &
          LHSvar %in% c("rho_base","c_u_ratio","c_e_ratio")) |
         (model=="instswitch+epc" &
          LHSvar %in% c("rho_base","c_u_ratio")) |
         (model=="pairform" &
          LHSvar %in% c("c_u_ratio","c_e_ratio"))))
## reorder factors
mL3$LHSvar <- factor(mL3$LHSvar,
                     levels=c("Beta1","Beta3",
                              "Duration1","Duration3",
                              "c_mean_base","c_e_ratio","rho_base",
                              "c_u_ratio"))
## FIXME: could also use labels argument to rename levels, labeller=label_parsed
## in facet_grid to get pretty math
ggplot(mL3,aes(LHSval,sumval,colour=model))+
    geom_point(pch=".",alpha=0.5)+
    facet_grid(sumvar~LHSvar,scales="free")+
    geom_smooth(se=FALSE)+labs(x="",y="")+
    scale_x_log10(expand=c(0,0.08))+  ## leave a little extra room?
    zero_margin
@

Finally, the sensitivity plots (Figure~\ref{fig:plot_sens}) show the effects 
of each parameter on the summary statistics. In almost every case the
effects of the parameters are monotonic; note that
the plot shows the effects of the \emph{unscaled} parameters, i.e.
before they have been adjusted to achieve a standard initial epidemic
growth rate.
The transmission rates ($\beta_P$, $\beta_D$)
and durations ($D_P$, $D_D$) in the primary and disease stages generally
decrease the equilibrium virulence, peak virulence, and peak time,
although the random and pair-formation+epc models have high, relatively
constant values with respect to these parameters. 
\todo{I know you explained to me before why beta and Duration go in
the same direction but I forget ...}
The base contact rate increases virulence and peak time in almost all
cases, although the pair-formation+epc model is again relatively insensitive.
The ratio of extra-pair to within-pair contact ($c_e/c_w$) affects
virulence in the instantaneous-switching+epc model, but not the pair-formation+epc
model (probably because the uncoupled individuals present in the pair-formation+epc
model make extra-pair contact by coupled individuals less important).
Surprisingly, neither of the pair-formation models is particularly sensitive to the 
rate of partnership formation ($\rho$); the rate of uncoupled contact
increases virulence and peak time in the pair-formation+epc model, 
which is the only model to which it applies\todo{is that right??}.

\section{Discussion}

All models must simplify the world.  Many constraints --- such as data
availability, computation time, or code complexity --- drive the need
for parsimony, with different constraints applying in different
contexts. The critical question that modelers must ask is whether the
simplified model gives adequate answers, or whether the
simplifications have led to qualitatively or quantitatively incorrect
conclusions.  This question is especially important for modelers who
are hoping that their conclusions will guide management decisions.

In the particular example of HIV virulence eco-evolutionary dynamics
and epidemiological, we reach the slightly ironic conclusion that the
effort put into building a more realistic model essentially cancels
out, leaving us in the same position as if we had ignored the problem
of epidemiological complexity entirely and used a naive random-mixing
contact model.  \cite{herbeck_hiv_2014} build a
network model of partnerships, but they set the ``partnership''
duration to 1 day --- very unrealistic in epidemic terms, but perhaps
actually more true to real-world HIV epidemiological dynamics than a
model with realistic partnership durations that neglects extra-pair
contact \citep{herbeck2016evolution}. Adding still further realism
might push the results farther away from the random-mixing model. For
example, our model forms partnerships randomly, and assumes that
extra-pair contact is randomly mixing across the population;
one could instead implement extra-pair contact as arising from
multiple concurrent partnerships (some, such as contact with sex
workers, of very short duration), or more structured partnership
formation (by age, ethnicity, or behaviour group). The effects of
other realistic complications such as explicit modeling of two
sexes (both in contact structure and differential transmission
probabilities), temporal and spatial variation in epidemic processes,
or presence of genetic variation in hosts are harder to predict.

Parameterization is one of the biggest challenges of epidemiological
modeling. In addition to following \champr\ by doing Latin hypercube
sampling across a wide range of epidemiological parameters, we 
calibrated each set of parameters to the same initial epidemic
growth rate, chosen to match the results of previous models
\citep{shirreff_transmission_2011}.  The previous models drew their
parameters from previous cohort studies from the 1980s
\todo{check dates!} \citep{hollingsworth_hiv1_2008}
rather than doing any explicit calibration to epidemic curves,
but they give reasonable order-of-magnitude
growth rates ($\approx 0.4~\textrm{year}^{-1}$) 
for the early stages of the HIV epidemic (although considerably
lower than estimates of $\approx 0.07-0.1~\textrm{year}^{-1}$
based on population genetic reconstructions \citep{faria_early_2014}).
\todo{do we know what $r$ Champredon et al. get??}
However, our reason for calibrating was not to match any
specific observed epidemic, but rather to make sure that
we were making meaningful comparisons across a range of
models with radically different epidemiological structures, and
hence involving different interpretations of the same quantitative
parameters.  For example, in models with instantaneous switching the
partnership dissolution rate $c$ is identical to the partnership
formation rate; in models with explicit partnership formation,
the partnership formation rate is also $c$ at equilibrium,
but might vary over the course of an epidemic.
It is not obvious whether models with equal parameters but
different structures should be directly compared; calibration
solves this problem.

More generally, any model that wants to be
taken seriously for management and forecasting purposes should
be calibrated to \emph{all} available data, using informative
priors to incorporate both realistic distributions of uncertainty
for all parameters from independent measurements \citep{elderd_uncertainty_2006}
and calibration from population-level observations of epidemic
trajectories. For practical (rather than exploratory) purposes,
such a procedure would also be an improvement on the common --- although not universal --- %
practice of assessing uncertainty over uniform ranges, rather than
distributions that allow more continuous variation in support over
the range of a parameter.

Researchers have documented that HIV virulence and set-point viral
load are changing, on time scales comparable to those portrayed here
(e.g., compare Figure~\ref{fig:virtraj} to \cite{herbeck_is_2012}'s
estimated rate of change of 1.3 \Lspvl\ per century [95\% CI -0.1 to
  3]), and have begun to build relatively realistic models that
attempt to describe how interventions such as mass antiretroviral
therapy (ART) can be expected to change the trajectory of virulence
evolution \citep{payne_impact_2014,herbeck2016evolution}.  While these
efforts are well-intentioned, we caution that epidemiological and
other structural details that are currently omitted from these models
could significantly change their conclusions.

\section{To do}
\begin{itemize}
\item clean up code and results
\item finish writing!
\item hard parts of the results
\begin{itemize}
\item explaining similarities among the models: why does implicit look most like pair formation model? (see R0comparison HTML)
\item explaining sensitivity plot
\end{itemize}
\item cite \cite{LipsitchNowak1995}
\end{itemize}

\begin{itemize}
\item fig 1. tweak size? remove redundant y axes/labels?
\item fig 2. 6-panel figure? log scale? boundary lines instead of ribbons for edges of envelopes? move labels away from axes?
\item fig 3. change order (flip instswitch and instswitch + epd)
\item fig 4. squash panels together?
\item fig 5. sensitivity plot (unscaled plot); check we havetake out plots that are not related ... implicit model doesn't use rho, cu or ce; random doesn't use rho, cmean?, cu, ce; instswitch doesn't have rho or cu or ce; models without epc don't have ce or cu; instswitch + epc doesn't have rho and cu (no uncoupled compartment). Maybe squash panels together, higher resolution
\end{itemize}

\appendix

See \href{http://tinyurl.com/bolker-jrsi-cites}{Bolker JRSI Google scholar cites} (\href{https://scholar.google.ca/scholar?oi=bibs&hl=en&cites=18202064200267310847}{alternative link})

\section{Model details}

Since we use multi-strain models, in which the distribution of \Lspvl\ has been discretized into a vector, we use a matrix notation to describe our models. The five states described in the Methods section are replaced with the following notations: $S$, $I_i$, $SS$, $SI_i$, $II_{ij}$, where the subscripts denote the strain that an individual is infected with. For example, $I_i$ is number of infected individuals with \Lspvl\ of $\alpha_i$, and $II_{ij}$ is the number of concordant, HIV-positive couples in which the two partners have \Lspvl\ of $\alpha_i$ and $\alpha_j$ (independent of order; $II_{ij}$ is synonymous with $II_{ji}$). 
Below, we use the Kronecker delta (i.e. $\delta_{ij}=1$ if $i=j$ and 1 otherwise) in a slightly non-standard fashion as an exponent, e.g. $2^{\delta_{ij}}$, to set a value to 2 when $i=j$ and 1 otherwise.

\subsection*{Models 1 and 2: partnership dynamics}

Single individuals acquire partners at rate $\rho$: $S' = - \rho S$ and $I_i' = - \rho I_i$. We follow Champredon et al's results and assume that single individuals are distributed into coupled states through binomial distribution\todo{? don't understand this wording (ditto below)}:
\todo{I think we need different notation for partnership formation (i.e. LHS is wrong)}

\begin{equation}
\begin{aligned}
(SS)' &= \frac{\rho S \cdot S}{2 (S + \sum_k I_k)}\\
(SI_i)' &= \frac{\rho S \cdot I_i}{S + \sum_k I_k}\\
(II_{ij})' &= \khalf \cdot \frac{\rho I_i \cdot I_j}{S + \sum_k I_k}
\end{aligned}
\end{equation}
In a single-strain model, both individuals in a concordant partnership, $II$, enter the single-infected compartment, $I$, when a partnership dissolves: $I' = 2 c II'$. However, this is not the case in a multi-strain model. If infected individuals in an $II$ partnership have different strains, they will enter different compartments; if they are both infected with the same strain, they both enter the same compartment. As described above, we use the Kronecker delta notation to implement this difference between the $i=j$ and $i\neq j$ cases.

Partnerships dissolve at rate $c$: $SS' = -c SS$, $SI_i' = -c SI_i$, and $II_{ij} = - c II_{ij}$. Unlike single strain model, where both individuals leaving the $II$ partnership would enter $I$, we have to account for strains which the individuals in concordant partnership are infected with (i.e. both partners in $II_{ii}$ enter $I_i$ whereas only one partner in $II_{ij}$ enters $I_i$).

\begin{equation}
\begin{aligned}
S' &= 2 c SS + \sum_k c SI_k \\
I_i' &= c SI_i + \sum_k 2^{\delta_{ik}} c II_{ik}\\
\end{aligned}
\end{equation}
Combining partnership formation and dissolution process yields the following equation:

\begin{equation}
\begin{aligned}
S' &= - \rho S + 2 c SS + \sum_k c SI_k \\
I_i' &= - \rho I_i + c SI_i + \sum_k 2^{\delta_{ik}} c II_{ik}\\
SS' &= \frac{\rho S \cdot S}{2 (S + \sum_k I_k)} - c SS\\
SI_i' &= \frac{\rho S \cdot I_i}{S + \sum_k I_k} - c SI_i\\
II_{ij}' &= \khalf \cdot \frac{\rho I_i \cdot I_j}{S + \sum_k I_k} - c II_{ij}
\end{aligned}
\end{equation}

\subsection*{Models 1 and 2: infection}

Within-couple transmission occurs in both models. An infected partner in $SI$ partnership transmits virus to a susceptible partner, and partnership state becomes $II$: $SI_i' = - \beta_i SI_i$. Since we assume that mutation occurs, $II_{ij}$, where $i \neq j$, can be formed from both $SI_i$ and $SI_j$ partnership: $II_{ij}' = M_{ij} \beta_i SI_i + M_{ji} \beta_j SI_j$. On the other hand, $II_{ii}$ can only be formed from an $SI_i$ partnership: $II_{ii}' = M_{ii} \beta_i SI_i$. Using the Kronecker delta notation, we obtain following set of equations that describe within-couple transmission dynamics:

\begin{equation}
\begin{aligned}
SI_i' &= - \beta_i SI_i\\
II_{ij} &=  \khalf \cdot (M_{ij} \beta_i SI_i + M_{ji} \beta_j SI_j)
\end{aligned}
\end{equation}

Champredon et al define the proportion of infectious extra-couple and uncoupled contact through the following term:

\begin{equation}
P = \frac{c_u I + c_e (SI + 2 II)}{c_u (S + I) + 2 c_e(SS + SI + II)}.
\end{equation}
Effective uncoupled, $c_u$, and extra couple, $c_e$, contact rate is the product of two terms: uncoupled/extra-couple contact rate $\times$ rate of transmission per contact. Therefore, the transmission rate per contact term in $c_u$ and $c_e$ is canceled out in the equation above. Using this property, we modify the equation above as follows:

\begin{equation}
P = \frac{r_u I + r_e (SI + 2 II)}{r_u (S + I) + 2 r_e(SS + SI + II)},
\end{equation}
where $r_u = c_u/c_w$ and $r_e = c_e/c_w$ are the relative uncoupled/extra-couple contact rates. This simplification is useful in a multi-strain model since we cannot multiply a vector with a single value (e.g. $c_u S$ in denominator) if we use Champredon et al's equation in its original form. Extending the above equation to multi-strain model so that $P_i$ represents the proportion of the extra-couple and uncoupled contact of an infected individual with strain $i$, we obtain:

\begin{equation}
P_i = \frac{r_u I_i + r_e (SI_i + \sum_k (II_{ik} + \delta_{ik} II_{ik}))}{r_u (S + \sum_k I_k) + r_e(2 SS + \sum_k 2 SI_k + \sum_l \sum_k (1 + \delta_{lk})II_{lk} )},
\end{equation}
Using the equation above, we can model uncoupled and extra-couple mixing. For convenience, uncoupled and extra-couple transmission rates, $c_u$ and $c_e$, will be replaced with $U_i = r_u \beta_i$ and $E_i = r_e \beta_i$ hereafter.

Single susceptible individuals become infected and enter single infected compartment at the total rate of $\sum_k P_k U_k S$. Through mutation, newly infected individuals are distributed into single infected compartments with different strains: $I_i' = \sum_k M_{ki} P_k U_k S$. Either partner in an $SS$ partnership can be infected, with the partnership state becoming $SI$, at the total rate of $\sum_i 2 P_i E_i SS$. The formation of $SI_i$ partnerships is similar to the process through which single susceptible individuals are distributed into single infected compartments: $SI_i' = \sum_k 2 M_{ki} P_k E_k SS$. Lastly, the susceptible partner in an $SI$ partnership can be infected due to uncoupled/extra-couple contacts and partnership can change to an $II$ partnership. As in the previous cases, $SI_i$ partnerships are lost at a rate of $\sum_k P_k E_k SI_i$. The mutation process is similar to that of within-couple transmission. The only difference is that the $\log_10$ SPVL of a newly infected partner is not determined by its social partner but from an extra-couple partner (i.e. the term $P_i$): $II_{ij}' = (1 -\frac{\delta_{ij}}{2})(\sum_k (M_{kj} P_k E_k SI_i + M_{ki} P_k E_k SI_j))$. Combining these equations we get the following set of equations that describe all the transmission dynamics \todo{something missing here?}

\subsection*{Model 1 and 2 - Disease induced mortality}

The disease induced death rate, $\lambda$, is given by taking the reciprocal of the total duration of the infection: $\lambda_i = 1/(D_A + D_P(\alpha_i) + D_D)$. Since we assume an SIS formulation, where infected individuals that die from infection are immediately replaced by an individual in the single susceptible compartment, we obtain the following equation for single infected individuals:

\begin{equation}
\begin{aligned}
S' &= \sum_k \lambda_k I_k \\
I_i &= - \lambda_i I_i \\
\end{aligned}
\end{equation}

If an infected individual in a partnership dies, the partnership dissolves. Thus, an $SI_i$ partnership dissolves at a rate $-\lambda_i$, and the susceptible partner enters the single susceptible compartment at rate $\lambda_i SI_i$ (due to the SIS formulation the the infected partner that dies also gives rise, at an equal rate, to an individual entering the single susceptible compartment):

\begin{equation}
\begin{aligned}
S' &= \sum_k 2 \lambda_k SI_k \\
SI_i &= - \lambda_i SI_i 
\end{aligned}
\end{equation}

Similarly, $II_{ij}$ partnerships dissolve at a rate $-(\lambda_i + \lambda_j)$, but two cases, when $i \neq j$ and $i = j$, must be considered separately. When an $II_{ij}$ partnership dissolves due to disease-induced mortality, where $i \neq j$, the death of the partner with strain $i$ causes its partner to enter $I_j$ compartment at rate $\lambda_j II_{ij}$, and vice versa. When an $II_ii$ partnership dissolves, the death of either partner causes the other partner to enter the $I_i$ compartment at rate $\lambda_i II_{ii}$, which sums up to $2\lambda_i II_{ii}$. Combining these dynamics yields:

\begin{equation}
\begin{aligned}
S' &= \sum_l \sum_k  2^{\delta_{lk}} \lambda_k II_{lk} \\
I_i' &=  \sum_k 2^{\delta_{ik}} \lambda_k II_{ik} \\
II_{ij}' &= -(\lambda_i + \lambda_j) II_{ij}
\end{aligned}
\end{equation}

Finally, combining all these equations give us the full model, which is Model 1. We can simply take out the uncoupled and extra couple transmission term to obtain equation 2\todo{?}:

\begin{equation}
\begin{aligned}
S' =& - \rho S + 2 c SS + \sum_k c SI_k - \sum_k P_k U_k S + \sum_k \lambda_k I_k \\
&+ \sum_k 2 \lambda_k SI_k + \sum_l \sum_k  2^{\delta_{lk}} \lambda_k II_{lk}\\
I_i' =&  - \rho I_i + c SI_i + \sum_k 2^{\delta_{ik}} c  II_{ik} + \sum_k M_{ki} P_k U_k S- \lambda_i I_i \\
&+ \sum_k 2^{\delta_{ik}} \lambda_k II_{ik} \\
SS' =& \frac{\rho S \cdot S}{2 (S + \sum_k I_k)} - c SS - \sum_i 2 P_i E_i SS \\
SI_i' =& \frac{\rho S \cdot I_i}{S + \sum_k I_k} - c SI_i - \beta_i SI_i + \sum_k 2 M_{ki} P_k E_k SS - \sum_k P_k E_k SI_i   \\
&- \lambda_i SI_i\\
II_{ij}' =& \khalf \cdot \frac{\rho I_i \cdot I_j}{(S + \sum_k I_k)} - c II_{ij} + \khalf \cdot (M_{ij} \beta_i SI_i + M_{ji} \beta_j SI_j) \\
&+ \khalf \cdot (\sum_k (M_{kj} P_k E_k SI_i + M_{ki} P_k E_k SI_j)) -(\lambda_i + \lambda_j) II_{ij}
\end{aligned}
\end{equation}

\subsection*{Models 3 and 4: Partnership dynamics}

Since model 3 and 4 assume instantaneous partnership formation, there are only three states: $SS$, $SI_i$, and $II_{ij}$. Partnership dissolution rates are equal to those of model 1 and 2: $SS' = -cSS$, $SI_i' = - cSI_i$, and $II_{ij}' = - II_{ij}$. Once individuals leave a partnership, they enter temporary compartments and are distributed into a partnership through binomial distribution:

\begin{equation}
\begin{aligned}
X &= 2 c SS + \sum_k c SI_k \\
Y_i &= c SI_i + \sum_k 2^{\delta_{ik}} c II_{ik} \\
SS' &= - c SS + \frac{X^2}{2 (X + \sum_k Y_k)}\\
SI_i' &= - c SI_i + \frac{X Y_i}{X + \sum_k Y_k}\\
II_{ij}' &= - c II_{ij} +\khalf \frac{Y_i Y_j}{X + \sum_k Y_k}.
\end{aligned}
\end{equation}

\subsection*{Model 3 and 4 - Infection}

Model 3 and 4 share the within-couple transmission term with model 1 and 2. Since there is no single state, only extra couple transmission exists:

\begin{equation}
P_i = \frac{r_e (SI_i + \sum_k (II_{ik} + \delta_{ik} II_{ik}))}{r_e(2 SS + \sum_k 2 SI_k + \sum_l \sum_k (II_{lk} + \delta_{kl} II_{lk}) )}.
\end{equation}
Movement from $SS$ state to $SI$ state and $SI$ to $SS$ is modeled through the same equation that is used in model 1 and 2.

\subsection*{Model 3 and 4 - Disease induced mortality}

Disease induced mortality is modeled similar to model 1 and 2. However, as single state does not exist in model 3 and 4, individuals that has left their partnerships due to death of their partners enter temporary compartments and form partners instantly:

\begin{equation}
\begin{aligned}
X &= \sum_k 2 \lambda_k SI_k + \sum_l \sum_k 2^{\delta_{lk}}  \lambda_k II_{lk} \\
Y_i &=  \sum_k  2^{\delta_{ik}}  \lambda_k II_{ik} \\
SS = &= \frac{X^2}{2 (X + \sum_k Y_k)}\\
SI_i &= - \lambda_i SI_i + \frac{X Y_i}{X + \sum_k Y_k}\\
II_{ij}' &= -(\lambda_i + \lambda_j) II_{ij} + \khalf \cdot \frac{Y_i Y_j}{X + \sum_k Y_k}
\end{aligned}
\end{equation}

Combining all these dynamics, we have equation 3. If we remove extra-couple transmission, we have equation 4.

\begin{equation}
\begin{aligned}
X =& 2 c SS + \sum_k c SI_k + \sum_k 2 \lambda_k SI_k + \sum_l \sum_k 2^{\delta_{lk}}  \lambda_k II_{lk}\\
Y_i =& c SI_i + \sum_k 2^{\delta_{ik}}  c II_{ik} + \sum_k  2^{\delta_{ik}}  \lambda_k II_{ik} \\
SS  =& - c SS + \frac{X^2}{2 (X + \sum_k Y_k)}  - \sum_i 2 P_i E_i SS\\
SI_i =& - c SI_i + \frac{X Y_i}{X + \sum_k Y_k} - \beta_i SI_i + \sum_k 2 M_{ki} P_k E_k SS\\
&- \sum_k P_k E_k SI_i - \lambda_i SI_i  \\
II_{ij}& - c II_{ij} +\khalf \frac{Y_i Y_j}{X + \sum_k Y_k} + \khalf \cdot (M_{ij} \beta_i SI_i + M_{ji} \beta_j SI_j) \\
&+ \khalf \cdot (\sum_k (M_{kj} P_k E_k SI_i + M_{ki} P_k E_k SI_j)) -(\lambda_i + \lambda_j) II_{ij}
\end{aligned}
\end{equation}

\subsection*{Model 5}

Following \cite{shirreff_transmission_2011}, Model 5 is an implicit instantaneous partnership formation model that uses an adjusted transmission rate, $\beta'$, that is derived from \cite{hollingsworth_hiv1_2008}'s approximate basic reproduction number:

\begin{equation}
\beta'_i = \frac{c \beta_i}{c + \beta_i + \lambda_i}.
\end{equation}
Thus, we get the following model:

\begin{equation}
\begin{aligned}
S' & = \sum_k \lambda_k I_k - \sum_k \beta'_k S I_k\\
I_i & = \sum_k M_{ki} \beta'_k S I_k - \lambda_i I_i
\end{aligned}
\end{equation}

\subsection*{Model 6}

Model 6 is a random mixing model. It is modeled in a same way as model 6 without the adjusted transmission rate:

\begin{equation}
\begin{aligned}
S' & = \sum_k \lambda_k I_k - \sum_k \beta_k S I_k\\
I_i & = \sum_k M_{ki} \beta_k S I_k - \lambda_i I_i
\end{aligned}
\end{equation}

\subsection*{Initial distribution of infected individuals}

We follow Champredon et al's result to calculate the initial distribution of infected individuals. For model 1 and 2, we have disease equilibrium state of $S^* = \frac{c}{c + \rho}$ and $SS^* = \frac{1-S^*}{2}$. We let $\epsilon = 10^{-4}$, which is the total number of infected individuals in the beginning of simulation and $D$ be the vector such that $D_i$ represent the proportion of individuals with \Lspvl\ of $i$. $Y_i$ is taken from normal distribution with mean 3 and is normalized so that $\sum_i D_i = 1$. Then, we have the following initial distribution of each states:

\begin{equation}
\begin{aligned}
S(0) &= (1 - \epsilon) S^* \\
SS(0) &= (1 - \epsilon)^2 SS^*\\
SI_i(0) &= 2 \epsilon (1-\epsilon) SS^* D_i\\
I_i(0) &=  \epsilon S^* D_i\\
II_{ij}(0) &=  \khalf 2\epsilon^2 SS^* D_i D_j.
\end{aligned}
\end{equation}
Since model 3 and 4 do not have single state, $SS^*=1$ at the disease free equilibrium and the initial distribution becomes as follows:

\begin{equation}
\begin{aligned}
SS(0) &= (1 - \epsilon)^2 SS^*\\
SI_i(0) &= 2 \epsilon (1-\epsilon) SS^* D_i\\
II_{ij}(0) &=  \khalf 2\epsilon^2 SS^* D_i D_j.
\end{aligned}
\end{equation}
Lastly, as model 5 is an implicit model, which does not consider different stages of partnership, we have the following initial distribution.

\begin{equation}
\begin{aligned}
S(0) &= 1 - \epsilon\\
I_i(0) &=  \epsilon D_i.
\end{aligned}
\end{equation}
Model 6 also shares the same distribution of initial infected individuals as model 5.

\bibliographystyle{chicago}
\bibliography{virulence}
\end{document}