### 2016-05-29 poster/hivlhs_eeid2016.Rnw
poster stuff

%\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}
\documentclass{beamer}
\usepackage{grffile}
\mode<presentation>{\usetheme{drexel}}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{amsmath,amsthm, amssymb, latexsym}
\usefonttheme[onlymath]{serif}

%% You can change the poster orientation (portrait vs. landscape) and the paper size (a0, a1, &c.) here:
\usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}

\usepackage{array,booktabs,tabularx}
\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns
\newcommand{\pphantom}{\textcolor{ta3aluminium}} % phantom introduces a vertical space in p formatted table columns??!!

%% MY MACROS %%
\newcommand{\R}{{\mathcal R}}
\newcommand{\Ro}{\mathcal{R}_0}
\newcommand{\Ihat}{{\hat{I}}}
\newcommand{\Imin}{I_{\rm min}}
\newcommand{\etal}{\emph{et al.}}
\newcommand{\fref}[1]{Figure~\ref{fig:#1}}

<<pkgs,warning=FALSE,message=FALSE, echo = FALSE>>=
library("knitr")
library("ggplot2"); theme_set(theme_bw(base_size = 32,
                                       base_family = "Times"))
opts_chunk$set(echo=FALSE,error=FALSE)
if (.Platform$OS.type=="windows") {
    windowsFonts(Times=windowsFont("Times"))
} 
zero_margin <- theme(panel.margin=grid::unit(0,"lines"))
zero_x_margin <-
    theme(panel.margin.x=grid::unit(0, "lines"))
library("MASS")
## use Dark2 rather than Set1 because colour #6 of Set1 is yellow (ugh/too light)
scale_colour_discrete <- function(...,palette="Dark2") scale_colour_brewer(...,palette=palette)
scale_fill_discrete <- function(...,palette="Dark2") scale_fill_brewer(...,palette=palette)
library("gridExtra")
library("plyr")  ## for ldply()
library("dplyr") ## for full_join()
library("reshape2")  ## for melt()
library("GGally") ## need BMB version
## devtools::install_github("bbolker/GGally")
## devtools::install_github("lionel-/ggstance")
library("ggstance")
@

<<load, echo = FALSE>>=
load("../simdata/combine_ev_LHS4.rda")
@

<<setup2,echo=FALSE>>=
orig_sum_labs <- c("peak_time","peak_vir","eq_vir","rel_peak")
new_sum_labs <- c("peak time","peak virulence","equilibrium virulence","relative peak")
fixfac2 <- function(x,atop=FALSE,newlines=FALSE) {
    if (atop) new_sum_labs <-
                  gsub("(.*) (.*)","atop(\\1,\\2)",new_sum_labs)
    factor(x,
           levels=orig_sum_labs,
           labels=new_sum_labs)
}
@

<<setup3, echo=FALSE,dev="tikz",cache=TRUE>>=
source("../R/hivFuns.R")
source("../R/LHSparms3.R")
Lspvl <- "$\\log_{10}~\\textrm{SPVL}$"
pars <- expand(ltab_mean)
dd <-with(pars,
     melt(data.frame(alpha=alpha,"transmission probability"=Beta2,
                     "duration (years)"=Duration2,check.names=FALSE),
          id.vars="alpha"))
gg_tradeoff <- ggplot(dd,aes(alpha,value))+facet_wrap(~variable,scale="free")+
    geom_line()+
    labs(x= Lspvl)
@

%%%\title{\huge On the Persistence of Invading Pathogens}
\title{\LARGE HIV virulence evolution in structured epidemic models}

\author{\href{http://www.math.mcmaster.ca/bolker/}{Ben Bolker and Sang Woo Park}}
\email{\href{mailto:bolker@mcmaster.ca}{\texttt{bolker@mcmaster.ca}}}
\institute[McMaster University]{McMaster University, Hamilton, Ontario, Canada %%\\\vspace{5mm}\texttt{earn@math.mcmaster.ca}
}
\date[June 2016]{June 2016}

%% Uncomment the following line to define a custom logo (or to tweak the size of the logo)
\DrexelLogo{\includegraphics[width=\WhenPortrait{1.75}{1}\linewidth]{images/Maclogo}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newlength{\columnheight}
\setlength{\columnheight}{105cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frame}
%\includegraphics[width=\textwidth]{images/photobanner.pdf}
\vspace{0.5cm}
  \noindent\begin{columns}
    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.33\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{block}{Summary}
\noindent
\hspace{-1.6cm}
Pathogens can evolve rapidly in response to changing
conditions (e.g., epidemic stage or public health interventions). 
Models of \textbf{eco-evolutionary dynamics} often
neglect important epidemiological processes, such as the formation and
dissolution of sexual partnerships. We
compared \textbf{models with a range of complexity} in their
representation of partnership dynamics and extra-partnership contact.
Simple random-mixing actually gave the closest match to the most realistic models; extra-pair contact washed out the effects
of epidemiological structure. Nevertheless, the large differences
in evolutionary dynamics among different epidemiological models
suggests that researchers trying to predict the evolution of
pathogens should proceed with caution.

\end{block}

\vspace{0.5cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{block}{Tradeoff theory}

<<tradeoff_fig,echo=FALSE,fig.width=10>>=
print(gg_tradeoff)
@ 
  \begin{itemize}
    \item \textbf{tradeoff theory}: virulence evolution
      mediated by transmission-vs-clearance tradeoff
    \item still debated \cite{EbertBull2003,alizon_adaptive_2015}
    \item HIV \cite{fraser_virulence_2014}: \textbf{set-point viral load}
      correlated with transmission probability, rate of progression to AIDS
    \item data from Rakai, Uganda
    \item eco-evolutionary virulence dynamics: \cite{shirreff_transmission_2011}
  \end{itemize}

\end{block}

%%\vfill
\vspace{0.5cm}

\begin{block}{Epidemiological structure}

  %\vspace{-3cm}
  \begin{center}
    \includegraphics[width=\textwidth]{images/champ1.png}
  \end{center}

Spectrum of model complexity: 
\begin{itemize}
\item \textbf{pair formation}: instantaneous or delayed?
\item \textbf{extra-pair contact} (\textbf{epc}): present or absent?
\item \textbf{implicit model}: no explicit partnerships,
  force of infection expression derived from $\Ro$ of pair-formation
  model (without epc)
\item \textbf{random-mixing model}: standard SIR model
\end{itemize}

Slightly simplified disease model (single stage only;
genetic but not phenotypic variation)

\end{block}

}
\end{minipage}
\end{beamercolorbox}
\end{column}
    % ---------------------------------------------------------%
    % end the column

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.33\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{block}{Basic evolutionary dynamics}

\vspace{-1cm}

<<virdyn,cache=TRUE,dev="tikz",fig.width=25,fig.height=30>>=
ss <- subset(sum_list[["vir_mat"]],tvec<750)
ss$model <- factor(ss$model, c("random",
         "pairform+epc", "pairform","instswitch+epc","instswitch","implicit"))

gg_virtraj <- ggplot(ss,
                     aes(tvec,mean,ymin=lwr,ymax=upr))+
    geom_line(aes(colour=model,linetype=model),lwd=1)+
    geom_ribbon(aes(fill = model), alpha=0.3)+
    labs(x="time (years)",y=Lspvl) +
    ## theme(axis.title.y = element_text(margin = margin(0,10,0,0)),
    ##  axis.title.x = element_text(margin = margin(10,0,0,0))) +
    facet_wrap(~model, nrow = 3) +
    zero_margin+
    theme(legend.text=element_text(size=16))
## scale_x_continuous(limits=c(0,2500))
print(gg_virtraj)
@

\vspace{0.5cm}

\end{block}            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{block}{Univariate summary}

<<univ,cache=TRUE,dev="tikz",fig.width=45,fig.height=30>>=
sL <- transform(sum_list[["sum_mat"]],rel_peak=peak_vir/eq_vir)
sL$model <- factor(sL$model, c("random","pairform+epc", "pairform", "instswitch+epc","instswitch" , "implicit"))
mL <- melt(sL,id.vars=c("model","run"))
## horrible hack (but doesn't help); subsample
## w <- which(mL$model=="random")
## mLw <- mL[-sample(w,size=length(w)*9/10,replace=FALSE),]
w <- with(mL,which(model=="random" & variable=="eq_vir"))
rval <- mean(mL$value[w])
mLw <- droplevels(mL[-w,])
mLw$variable <- fixfac2(mLw$variable)

gg_univ <- ggplot(mLw,aes(value,model,fill=model))+
	geom_violinh(width=1)+
	facet_wrap(~variable,scale="free_x")+
	guides(fill=guide_legend(reverse=TRUE))+
	labs(y="")+
	geom_point(data=data.frame(model="random",variable="equilibrium virulence",
                               value=rval),pch=22,size=3,show.legend=FALSE) +
	zero_x_margin
print(gg_univ)
@

\end{block}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{0.5cm}

\begin{block}{Bivariate summary}

<<ggfuns,echo=FALSE>>=
trim_gg <- function(gg,hack_spaces=TRUE) {
    n <- gg$nrow
    gg$nrow <- gg$ncol <- n-1
    v <- 1:n^2
    gg$plots <- gg$plots[v>n & v%%n!=0]
    gg$xAxisLabels <- gg$xAxisLabels[-n]
    gg$yAxisLabels <- gg$yAxisLabels[-1]
    ## ggpairs tries to parse labels, so this fails if they
    ## have spaces in them ... maybe there's a better way?
    if (hack_spaces) {
        gg$xAxisLabels <- gsub("_"," ",gg$xAxisLabels)
        gg$yAxisLabels <- gsub("_"," ",gg$yAxisLabels)
    }
    return(gg)
}

tweak_legends_gg <- function(gg,pos=1.5) {
    n <- gg$nrow
    for (i in 1:n) {
        for (j in 1:n) {
            inner <- getPlot(gg,i,j)
            if (i==1 & j==1) {
                inner <- inner + ggplot2::theme(legend.position=c(pos,0.5)) +
                    guides(colour = guide_legend(override.aes = list(size=6)))
            } else {
                inner <- inner + ggplot2::theme(legend.position="none")
            }
            gg <- putPlot(gg,inner,i,j)
        }
    }
    return(gg)
}

tweak_colours_gg <- function(gg) {
    n <- gg$nrow
    for (i in 1:n) {
        for (j in 1:n) {
            inner <- getPlot(gg,i,j)
            inner <- inner+scale_colour_brewer(palette = 'Dark2')
            gg <- putPlot(gg,inner,i,j)
        }
    }
    return(gg)
}
@

<<ggpairs,echo=FALSE,fig.width=25,fig.height=25>>=
sL2 <- sL
names(sL2)[na.omit(match(orig_sum_labs,names(sL)))] <- gsub(" ","_",new_sum_labs)
    ## paste0("`",new_sum_labs,"`")
ggp1 <- ggpairs(sL2,
        mapping = ggplot2::aes(color = model),
        columns=3:6,
        legends=TRUE,
        lower = list(continuous = wrap("points",size=0.1)), ## alpha = 0.3,size=0.5)),
        diag = list(continuous = "blankDiag"),
        upper = list(continuous = "blank"))
## http://stackoverflow.com/questions/14711550/is-there-a-way-to-change-the-color-palette-for-ggallyggpairs-using-ggplot
## have to change plot one panel at a time
ggp2 <- tweak_colours_gg(tweak_legends_gg(trim_gg(ggp1)))
## add 1-to-1 line
inner <- getPlot(ggp2,2,1)
inner <- inner+geom_abline(intercept=0,slope=1,lty=2)
ggp2 <- putPlot(ggp2,inner,2,1)
## tweak scale
inner <- getPlot(ggp2,3,1)
inner <- inner+scale_x_continuous(breaks=seq(2.5,4.5,by=0.5))
ggp2 <- putPlot(ggp2,inner,3,1)
print(ggp2,spacingProportion=0)
@   
\end{block}

}
          % ---------------------------------------------------------%
          % end the column
        \end{minipage}
      \end{beamercolorbox}
   \end{column}

    % ---------------------------------------------------------%
    % end the column
    \begin{column}{.33\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content

            \begin{block}{Sensivity}
<<sens,dev="png",echo=FALSE>>=
mL3 <- subset(mL2,
       !((model=="implicit" &
          LHSvar %in% c("rho_base","c_u_ratio","c_e_ratio")) |
         (model=="random" &
          LHSvar %in% c("rho_base","c_u_ratio","c_e_ratio")) |
         (model=="instswitch" &
          LHSvar %in% c("rho_base","c_u_ratio","c_e_ratio")) |
         (model=="instswitch+epc" &
          LHSvar %in% c("rho_base","c_u_ratio")) |
         (model=="pairform" &
          LHSvar %in% c("c_u_ratio","c_e_ratio"))))
## reorder factors
mL3$LHSvar <- factor(mL3$LHSvar,
                     levels=c("Beta1","Beta3",
                              "Duration1","Duration3",
                              "c_mean_base","c_e_ratio","rho_base",
                              "c_u_ratio"),
                     labels = c("beta[P]", "beta[D]", "D[P]", "D[D]",
                                "c", "c[e]/c[w]", "rho", "c[u]/c[w]"))

## FIXME: could also use labels argument to rename levels, labeller=label_parsed
## in facet_grid to get pretty math

mL3$model <- factor(mL3$model, c("random","pairform+epc", "pairform", "instswitch+epc","instswitch" , "implicit"))

mL3$sumvar <- fixfac2(mL3$sumvar)

L <- function(labels,multi_line=TRUE) {
    r <- if (all(grepl(" ",labels[[1]]))) {
             list(as.character(gsub(" ","\n",labels[[1]])))
    } else {
        label_parsed(labels,multi_line=multi_line)
    }
    return(r)
}

brkfun <- function(x) {
    nx <- 1
    n <- 2
    brks <- axisTicks(log(x), log=TRUE,  nint = n)
    ## browser()
    brks <- brks[brks>x[1] & brks<x[2]]
    nx <- length(brks)
    ## cat("brk1",brks,"\n")
    while (nx<2 && n<10) {
        n <- n+1
        ## cat(n,"\n")
        brks <- axisTicks(log10(x), log=TRUE,  nint = n)
        brks <- brks[brks>x[1] & brks<x[2]]
        ## cat("brk2",brks,"\n")
        nx <- length(brks)
    }
    if (n==10) browser()
    r <- brks[c(1,length(brks))]
    ## hacks
    if (all(r==c(0.01,1))) r <- c(0.02,0.5)
    if (all(r==c(0.2,5))) r <- c(0.5,5)
    return(r)
}

ggplot(mL3,aes(LHSval,sumval,colour=model))+
    geom_point(pch=".",alpha=0.5)+
    facet_grid(sumvar~LHSvar,scales="free",
               labeller = L)+
    geom_smooth(se=FALSE)+labs(x="",y="")+
    ## leave a little extra room?
    scale_x_log10(expand=c(0,0.08),breaks=brkfun)+
    zero_margin
@               
              \end{block}
            \begin{block}{Conclusions}
\end{block}
\begin{block}{References}

{\tiny
%% Add numbered references
\setbeamertemplate{bibliography item}[text]
%% Stop line breaks
\setbeamertemplate{bibliography entry title}{}
\setbeamertemplate{bibliography entry location}{}
\setbeamertemplate{bibliography entry note}{}
%% Stop font color changes
% \setbeamercolor*{bibliography item}{fg=black}
% \setbeamercolor*{bibliography entry title}{fg=black}
% \setbeamercolor*{bibliography entry author}{fg=black}
% \setbeamercolor*{bibliography entry location}{fg=black}
% \setbeamercolor*{bibliography entry note}{fg=black} 
%% Generate bibliography
\bibliographystyle{vancouver}
\bibliography{../ms/virulence}
}

\end{block}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{0.5cm}

\begin{block}{Acknowledgements}

We thank Christophe Fraser for providing access to code, and 
the Natural Sciences and Engineering Research Council of Canada (NSERC)
for funding.

\end{block}

          }
        \end{minipage}
      \end{beamercolorbox}
   \end{column}

 \end{columns}
\end{frame}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End: