### 2016-09-12 misc/fullModel.rmd
add full model with risk het stuff

---
title: "Attempting to write a full model"
output: html_document
---

I want to write a full model. Let's start by considering partnership dynamics:

$$
\begin{aligned}
N_i &= - \rho_i N_i + \sum_k 2^{\delta_{ik}} NN_{ik} \frac{c_i + c_k}{2}\\
NN_{ij} &= \frac{1}{2^{\delta_{ij}}} \frac{\rho_i N_i \rho_j N_j}{\sum_k \rho_k N_k} - \frac{c_i + c_j}{2} NN_{ij}
\end{aligned}
$$

Let's put that into a code:
```{r load, echo = FALSE, message=FALSE}
library("deSolve")
library("reshape2")
library("ggplot2"); theme_set(theme_bw())
source("../simFuns.R")  ## for transform.list
source("../hivFuns.R")
source("../Param.R")
source("../hivModels.R")
```
```{r}
calc_yini_partner <- function(parameters){
  with(c(expand(parameters)),{
    yini <- list(
      N  = p.risk,
      NN = matrix(0, n.risk, n.risk))
    return(yini)
  })
}

model1 <- function(parameters) {
  pp <- expand(parameters)
  attach(pp)

  cc_mat <- outer(r.risk * c_mean, r.risk * c_mean, "+")/2

  detach(pp)

  g <- function(t,yini,parameters) {
      with(as.list(c(yini,pp)), 
      {
          N <- yini[1:n.risk]
          NN <- matrix(yini[-(1:n.risk)], n.risk, n.risk)
          NN_adj <- NN
          diag(NN_adj) <- 2 * diag(NN_adj)

          frate <- r.risk * rho * N
          f.sum <- sum(frate)

          dN <- - frate + colSums(cc_mat * NN_adj)

          f.NN <- outer(frate, frate, "*")/f.sum 
          diag(f.NN) <- diag(f.NN)/2

          dNN <- f.NN - cc_mat * NN

          list(c(dN, dNN))
      })
  }
}
```

This model seems to be running fine. Now, I would like to calculate the single/couple ratios at equilibrium.... It seems too complicated especially with infected individuals... I can probably get away with it by deriving it numerically:

```{r}

calc_yini_full <- function(parameters){
    with(c(expand(parameters)),{
        c = Ini_I
        I_vec = c * I_initialize(parameters)
        yini <- list(
            S  = (1 - c) * p.risk,
            I = c(outer(I_vec, p.risk)),
            SS = matrix(0, n.risk, n.risk),
            SI = matrix(0, n.risk, n.risk * n.alpha),
            II = matrix(0, n.risk * n.alpha, n.risk *  n.alpha))
        return(yini)
    })
}

colSums2 <- function(data, by){
    if(is.vector(data)){
        data <- matrix(data, nrow = 1)
    }
    length <- ncol(data)
    n <- length/by
    z <- colSums(data)
    z.mat <- matrix(z, ncol = n)
    return(colSums(z.mat))
}

makeMutMat <- function(xpars) {
    p <- with(xpars, {
        p <- matrix(NA, n.alpha, n.alpha)
        for(i in 1:n.alpha){
            x <- min.alpha + (i-1)*d.alpha
            denom <- 
                diff(pnorm(c(min.alpha-d.alpha/2,max.alpha+d.alpha/2),
                    mean=x, sd=Vm))
            for (j in 1:n.alpha){
                k <- min.alpha + (j-1)*d.alpha
                p[i,j] = diff(pnorm(k + c(-1,1)*d.alpha/2, mean = x, sd = Vm))/denom
            }
        }
    p
    })
    return(p)
}

model2 <- function(parameters,
                   infection = FALSE,
                   mortality = FALSE) {
    pp <- expand(parameters)
    attach(pp)

    S.ind <- 1:n.risk
    I.ind <- (n.risk+1):(n.risk+n.risk*n.alpha)
    SS.ind <- (n.risk+n.risk*n.alpha+1):(n.risk+n.risk*n.alpha+n.risk*n.risk)
    SI.ind <- (n.risk+n.risk*n.alpha+n.risk*n.risk+1):(n.risk+n.risk*n.alpha+n.risk*n.risk+n.alpha*n.risk*n.risk)
    II.ind <- (n.risk+n.risk*n.alpha+n.risk*n.risk+n.alpha*n.risk*n.risk+1):(n.risk+n.risk*n.alpha+n.risk*n.risk+n.alpha*n.risk*n.risk+n.alpha*n.alpha*n.risk*n.risk)

    rho2 <- r.risk * rho
    c2 <- r.risk * c_mean

    rho2.I <- rep(rho2, each = n.alpha)

    cc_mat <- outer(c2, c2, "+")/2

    c2.I <- rep(c2, each = n.alpha)
    cc_mat.I <- outer(c2, c2.I, "+")/2

    cc_mat.II <- outer(c2.I, c2.I, "+")/2

    alpha2 <- rep(alpha, n.risk)

    #infection
    if(infection){
        p <- makeMutMat(parameters)

        beta.SI <- matrix(rep(Beta, n.risk*n.risk), nrow = n.risk, byrow = TRUE)

        distribute.SIinf <- function(SIinf){
            attach(pp)
            dist.SI <- matrix(0, n.risk * n.alpha, n.risk * n.alpha)
            for(i in 1:n.risk){
                vec.i <- c(1:n.alpha) + n.alpha * (i-1)
                for(j in 1:n.risk){
                    vec.j <- c(1:n.alpha) + n.alpha * (j-1)
                    dist.SI[vec.i,vec.j] <- SIinf[i,vec.j] * p
                }
            }
            detach(pp)
            return(dist.SI)
        }

        c_u_ratio2 <- r.risk * c_u_ratio
        c_e_ratio2 <- r.risk * c_e_ratio

        c_u_ratio2.I <- rep(c_u_ratio2, each = n.alpha)
        c_e_ratio2.I <- rep(c_e_ratio2, each = n.alpha)

        c_u2 <- outer(r.risk, c_u)

        r.risk.couple <- outer(r.risk, r.risk, "+")

    }

    if(mortality){
        lam2 <- rep(lam, n.risk)
        lammat_dis2 <- matrix(rep(lam2, n.alpha*n.risk), n.alpha * n.risk, n.alpha * n.risk)
        diag(lammat_dis2) <- 2 * diag(lammat_dis2)
        lammat_adj2 <- outer(lam2, lam2, "+")

    }

    inf.SI.from <- matrix(0, n.risk, n.risk * n.alpha)
    inf.SI.to <- matrix(0, n.risk * n.alpha, n.risk *  n.alpha)
    inf.S.from <- rep(0, n.risk)
    inf.S.to <- rep(0, n.risk * n.alpha)
    inf.SS.from <- matrix(0, n.risk, n.risk)
    inf.SS.to <- matrix(0, n.risk, n.risk * n.alpha)
    inf.SI.from2 <- matrix(0, n.risk, n.risk * n.alpha)
    inf.SI.to2 <- matrix(0, n.risk * n.alpha, n.risk *  n.alpha)

    mort.I.out <- rep(0, n.risk * n.alpha)
    mort.I.sus <- rep(0, n.risk)
    mort.SI.out <- matrix(0, n.risk, n.risk * n.alpha)
    mort.SI.sus <- rep(0, n.risk)
    mort.II.out <- matrix(0, n.risk * n.alpha, n.risk *  n.alpha)
    mort.II.I <- rep(0, n.risk * n.alpha)
    mort.II.sus <- rep(0, n.risk)

    detach(pp)

    g <- function(t,yini,parameters) {
        with(as.list(c(yini,pp)), 
        {

            S <- yini[S.ind]
            I <- yini[I.ind]
            SS <- matrix(yini[SS.ind], n.risk, n.risk)
            SI <- matrix(yini[SI.ind], n.risk, n.risk * n.alpha)
            II <- matrix(yini[II.ind], n.risk * n.alpha, n.risk * n.alpha)

            ##Partnership formation rate
            N <- S + colSums2(I, n.alpha)

            frate <- rho2 * N
            f.sum <- sum(frate)

            SS_adj <- SS
            diag(SS_adj) <- 2 * diag(SS_adj)

            f.SS <- outer(rho2 * S, rho2 * S)/f.sum
            diag(f.SS) <- diag(f.SS)/2

            f.SI <- outer(rho2 * S, rho2.I * I)/f.sum

            II_adj <- II
            diag(II_adj) <- 2 * diag(II_adj)

            f.II <- outer(rho2.I * I, rho2.I * I)/f.sum
            diag(f.II) <- diag(f.II)/2

            NN <- colSums(SS_adj) + rowSums(SI) + colSums2(SI, n.alpha) + colSums2(II_adj, n.alpha)

            ##Partnership dissolution rate
            d.SS <- cc_mat * SS

            d.SI <- cc_mat.I * SI

            ##trying to model disease dynamics

            if(infection){
                ##within couple infection
                inf.SI.from <- beta.SI * SI
                inf.SI.tmp <- distribute.SIinf(inf.SI.from)
                inf.SI.to <- inf.SI.tmp + t(inf.SI.tmp)
                diag(inf.SI.to) <- diag(inf.SI.to)/2

                ##extra couple infection
                #prop <- (c_u_ratio * I + c_e_ratio * (SI + colSums(II_adj)))/
                #(c_u_ratio * (S + sum(I)) +
                #c_e_ratio * (2 * SS + 2 * sum(SI) + sum(II_adj)))

                prop.mat <- matrix(c(c_u_ratio2.I * I + c_e_ratio2.I * (colSums(SI) + colSums(II_adj))), ncol = n.risk)
                prop.num <- rowSums(prop.mat)
                prop.denom <- sum(c_u_ratio2 * N + c_e_ratio2 * NN)
                prop <- prop.num/prop.denom

                inf.S.rate <- S * sweep(c_u2, 2, prop, "*")
                inf.S.from <- rowSums(inf.S.rate)
                inf.S.to <- c(apply(inf.S.rate, 1, function(x) x %*% p))

                inf.ce.tmp <- c_e * prop
                risk.SS <- r.risk.couple * SS
                inf.SS.from <- risk.SS * sum(inf.ce.tmp)

                ce.mut <- inf.ce.tmp %*% p
                inf.SS.tmp2 <- outer(c(t(r.risk * SS_adj)), c(ce.mut))
                inf.SS.to <- matrix(c(t(inf.SS.tmp2)), nrow = n.risk, byrow = TRUE)

                SI2 <- r.risk * SI
                inf.SI.from2 <- sum(inf.ce.tmp) * SI2

                inf.SI.to2 <- matrix(0, n.risk * n.alpha, n.risk * n.alpha)

                for(i in 1:n.risk){
                    tmp.vec <- c(1:n.alpha) + (i-1) * n.alpha
                    inf.SI.to2[,tmp.vec] <- outer(SI2[i,], ce.mut)
                }

                inf.SI.to2 <- inf.SI.to2 + t(inf.SI.to2)
                diag(inf.SI.to2) <- diag(inf.SI.to2)/2
            }

            if(mortality){
                mort.I.out <- lam2 * I
                mort.I.sus <- colSums2(mort.I.out, n.alpha)

                mort.SI.out <- sweep(SI, 2, lam2, "*")
                mort.SI.sus <- rowSums(mort.SI.out) + colSums2(mort.SI.out, n.alpha)

                mort.II.out <- lammat_adj2 * II
                mort.II.I <- colSums(lammat_dis2 * II)
                mort.II.sus <- colSums2(t(lammat_dis2 * II), n.alpha)
            }

            dS <- - rho2 * S + colSums(cc_mat * SS_adj) + rowSums(d.SI) - inf.S.from + mort.I.sus +
                mort.SI.sus + mort.II.sus
            dI <- - rho2.I * I + colSums(d.SI) + colSums(cc_mat.II * II_adj) + inf.S.to - mort.I.out + mort.II.I
            dSS <- f.SS - d.SS - inf.SS.from
            dSI <- f.SI - d.SI - inf.SI.from + inf.SS.to - inf.SI.from2 - mort.SI.out
            dII <- f.II - cc_mat.II * II + inf.SI.to + inf.SI.to2 - mort.II.out

            ##total ratio of people in the risk group
            #tot <- N + NN

            ##total number of couples
            #coup <- sum(SS[lower.tri(SS, diag = TRUE)]) + sum(SI) + sum(II[lower.tri(II, diag = TRUE)])

            ##total susceptible
            #sus <- S + colSums(SS_adj) + rowSums(SI)
            inf <- I + colSums(SI) + colSums(II_adj)

            tot_I <- sum(inf)
            tot_V <- sum(alpha *  inf)

            test <- inf.S.to

            list(c(dS, dI, dSS, dSI, dII), tot_I = tot_I, mean_V = tot_V/tot_I)
        })
    }
}

```

OK I think I have the full model... Now I need to test this somehow...

```{r comparison}
tvec <- seq(1,100, 0.1)
yini1 <- calc_yini_partner(HIVpars.mean)
yini2 <- calc_yini_full(HIVpars.mean)
r1 <- lsoda(unlist(yini1), func = model1(HIVpars.mean), parms = HIVpars.mean, times = tvec)
r2 <- lsoda(unlist(yini2), func = model2(HIVpars.mean), parms = HIVpars.mean, times = tvec)

N.risk1 <- with(HIVpars.mean, return(r1[,2:(1+n.risk)]))
N.risk2 <- with(expand(HIVpars.mean),{
        S <- r2[,2:(1+n.risk)]
        I <- apply(r2[,(2+n.risk):(1+n.risk+n.risk*n.alpha)], 1, function(x) colSums2(x, n.alpha))
        return(S + t(I))
})

all.equal(unname(N.risk1), unname(N.risk2))

matplot(N.risk1, type = "l")
matplot(N.risk2, type = "l")
```

It seems like partnership dynamics work fine. Can we check disease dynamics?

```{r disease}
pars <- transform(HIVpars.mean, n.risk = 1, r.risk = 1, p.risk = 1, scale_all = 5)

calc_yini_tmp <- function(parameters){
  with(c(expand(parameters)),{
    c = Ini_I
    I_vec = c * I_initialize(parameters)
    yini <- list(
      S  = (1 - c),
      SS = 0,
      SI = rep(0, n.alpha),
      I = I_vec,
      II = matrix(0, n.alpha, n.alpha))
    return(yini)
  })
}

yini_norisk <- calc_yini_tmp(pars)
yini_full <- calc_yini_full(pars)

tvec <- seq(1, 30, 0.1)
r_norisk <- lsoda(unlist(yini_norisk), func = gfun(pars), parms = pars, times = tvec)
r_full <- lsoda(unlist(yini_full), func = model2(pars, infection = TRUE, mortality = TRUE), parms = pars, times = tvec)

all.equal(r_norisk[,488],r_full[,488])

matplot(cbind(r_norisk[,488], r_full[,488]), type = "l")
matplot(cbind(r_norisk[,487], r_full[,487]), type = "l")
```

Introducing heterogeneity:

```{r}
pars2 <- transform(HIVpars.mean, n.risk = 2, r.risk = c(0.8,4.8), p.risk = c(0.95, 0.05), scale_all = 5)
yini_het <- calc_yini_full(pars2)
r_het <- lsoda(unlist(yini_het), func = model2(pars2, infection = TRUE, mortality = TRUE), parms = pars2, times = tvec)

matplot(cbind(r_full[,488], r_het[,1899]), type = "l")
matplot(cbind(r_full[,487], r_het[,1898]), type = "l")
```

One more test. This should be identical to homogeneous model

```{r}
pars3 <- transform(HIVpars.mean, n.risk = 2, r.risk = c(1,1), p.risk = c(0.5, 0.5), scale_all = 5)
yini_test <- calc_yini_full(pars3)
r_test <- lsoda(unlist(yini_test), func = model2(pars3, infection = TRUE, mortality = TRUE), parms = pars3, times = tvec)

matplot(cbind(r_full[,488], r_test[,1899]), type = "l")
matplot(cbind(r_full[,487], r_test[,1898]), type = "l")
```

Seems like this model works?