### 2016-10-19 ms/plosHIVcover_resubm.tex

\documentclass[10pt]{letter}

\usepackage{graphicx}
\usepackage{myltr2}
\usepackage{times}
%\usepackage[sort&compress]{natbib}

\bibliographystyle{abbrv}

\makeatletter
% http://tex.stackexchange.com/questions/18033/using-bibtex-with-letter-class
\newenvironment{thebibliography}[1]
     {\list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
\newcommand\newblock{\hskip .11em\@plus.33em\@minus.07em}
\makeatother

\newcommand{\revcomment}[1]{\emph{#1}}
\newcommand{\response}[1]{#1}

\begin{document}

\date{\today}

\signature{\includegraphics[height=0.35in]{bbsig3.png}\\Benjamin Bolker}

\begin{letter}{
}

\opening{To the editor,}

\revcomment{The abstract really needs to be improved. It’s missing the main points
of the paper (that different contact patterns change the predicted
trajectories and optimums of virulence), and has unnecessary minor
details (like the fact that parameters were sampled using a Latin
hypercube).}

\response{We've followed the reviewer's suggestions in adding
important points and removing unimportant details from the abstract}

\revcomment{The introduction is really lacking a cogent description of the study of
the evolution of virulence, and the motivation of the paper would be
totally unclear to someone not already very familiar with this field.
The main points that are needed to spell out this issue are:

[... detailed suggestions omitted]
}

\response{We thank the reviewer for pointing this out --- we've 
been over this ground so often that we've come to take this basic
framework for granted!  We've 
}
The paper claims to add considerations of population structure to the
study of virulence but only focuses on the rules of partnership
formation, while assuming all individuals have identical behavior
patterns, and not on another major aspect of population structure, which
is the heterogeneity in the number of partners (or rates of partnership
change) among individuals. In the case of a static network, this would
be considered the degree heterogeneity. Many studies have shown that
sexual contact networks tend to me more on the scale-free spectrum, and
so the assumption of uniformity of behavior seems weak. This study would
be most impactful if on top of examining these switch patterns it
examined the influence of contact heterogeneity.

In Figure 3 it would be helpful to also include time to progression to
AIDS as a figure panel

Reviewer #2: Modeling studies have shown that HIV could evolve towards
immediate virulence where the trade-off between host survival and
transmission is optimized. The aim of this study is to investigate how
different assumptions about sexual partnership dynamics affect the
transient evolution of HIV virulence in a population. While this is a
clear objective, the study is rather technical and, in my opinion, does
not provide major insights or conclusions. My major concerns regarding
the approach of the study is given below. Furthermore, there are several
issues in terms of structure and the model description that should be
considered by the authors.

Approach:
The authors argue that they compare more “realistic” models (based on
ref. 14) to the relatively simple model by Shireff et al. (ref. 10). In
my view, all these models are quite simplistic in that they do not take
into account age-structure and heterogeneity in sexual behavior.
Essentially, all these models are random mixing models where every
individual can make contact with every other individual. The only
difference to the random mixing model by Shireff et al. is that
transmission can occur in ongoing partnerships and that some models
allow for concurrency. Arguably much more important aspects that affect
HIV epidemics and virulence evolution are the above-mentioned age- and
risk-structure. Furthermore, the authors use a single-stage disease
model and do not distinguish between differences in transmissibility
between the acute (primary) and chronic (asymptomatic) stage of
infections, which might influence HIV virulence evolution more than
differences in the sexual partnership dynamics. Hence, I am not
convinced that the authors use a “more realistic” model than the one by
vShireff et al.

Structure:
- The references to the equations are not always correct.
- Figure 1 is described in Materials and Methods before the model is
fully explained. Consequently, the parameter r cannot be understood.
- Tables and Figures often lack dimensions, i.e., is is unclear what the
axes represent. Table 1 is more or less useless without providing
dimensions for the various parameters.
- The author summary seems to provide more information about the results
than the abstract. However, the authors use SPVL and progression to AIDS
as two different proxies for virulence which can be confusing.

Model:
The authors make use of a previously published model from ref. 14. This
model is based on the pair-formation formalism initially developed by
Dietz & Hadeler (1988, J Math Biol, PMID: 3351391) and later used by
Kretzschmar et al. (1994 & 1998, Math Biosci, PMID: 7833594 & 9597826).
I feel that the authors should refer to some of these original
publications as well. Unfortunately, the description of the models in
appendix S1 seems to be partially erroneous or incorrect. For example,
the primes for the derivatives are often missing (e.g., equation 22, 24
and 25) or it is unclear whether equations represent state variables or
derivates thereof. Also, the reasoning behind the “instswitch” models is
unclear to me. The authors say that “Once individuals leave a
partnership, they enter temporary compartments…”. The variables (or
derivatives?) X and Y_i in equation 19 consist of positive terms only,
so how do individuals leave these compartments? In my understanding the
standard pair model (equation 9) can easily account for instantaneous
partnership formation by simply setting the pair formation rate rho to a
high value.
 ------------------------------------------------------------------------

*Reviewers who wish to reveal their identities to the authors and other
reviewers should include their name here (optional). These names will
not be published with the manuscript, should it be accepted.*

Reviewer #1: (No Response)

Reviewer #2: (No Response)

\closing{Sincerely,}

\bibliography{virulence}

\end{letter}
\end{document}