### 2017-01-13 figures/diagram_geom.tex
attempt at tikz pictures!

\tikzset{
	auto,
	compartment/.style={
		rectangle, minimum size=9mm, rounded corners=2mm,
		thick, draw=black!15, top color=white,bottom color=black!30
	},
	%
	bigcompartment/.style={
		rectangle, minimum size=24mm, rounded corners=2mm,
		thick, draw=black!15, top color=white,bottom color=black!20
	},
	%
	point/.style={
		circle, inner sep=2pt, fill=black!5
	},
	%
	mytextbox/.style={
		rectangle, text=black!50, thin, 
		draw=white, top color=white,bottom color=white, fill=white
	}
}
% http://tex.stackexchange.com/questions/40388/avoid-arrow-crossing-over-blocks-in-flow-chart
\tikzstyle{line} = [draw, -latex']