### 2017-01-13 figures/instswitch.tex

\newcommand{\instswitchflowvec}{ 
	\draw[->, very thick] (SS) to (SI); 
	\draw[->, very thick] (SI) to (II); 
}

\newcommand{\instswitchnoepcinf}{
	\draw[->, blue, dashed, very thick] (SI)
		to [bend right=45] (SItoII);
	\draw[->, blue, dashed, very thick] (SI)
		to [bend left=45] (SStoSI);

}

\newcommand{\instswitchboxes}{
	\node(SS)[bigcompartment]{SS};
	\node(SStoSI)[right=of SS]{};
	\node(SI)[bigcompartment, right=of SStoSI]{SI};
	\node(SItoII)[right=of SI]{};
	\node(II)[bigcompartment, right=of SItoII]{II};
}

\newcommand{\pairformboxes}{
        \node(S)[bigcompartment]{S};
        \node(StoI)[right= of S]{};
        \node(I)[bigcompartment,right= of StoI]{I};
	\node(SS)[bigcompartment,below left= of S]{SS};
	\node(SStoSI)[right=of SS]{};
	\node(SI)[bigcompartment, right=of SStoSI]{SI};
	\node(SItoII)[right=of SI]{};
	\node(II)[bigcompartment, right=of SItoII]{II};
}